{{< include _setup.qmd >}}

# Statistical tests

```{r}
# This script runs balance tests comparing the universe of residential parcels
# to the sample of residential parcels with sales in the previous 8 years
# that are used to train the data department's CAMA.

# GATHER DATA ----

# Vars to test across
vars <-
  grep("char_|3435|acs5_median|township_name",
    names(chars_data),
    value = TRUE
  ) %>%
  grep("percentile|_apts|_ncu|_use|_qlty",
    .,
    value = TRUE,
    invert = TRUE
  )

# All residential parcel characteristics from time frame, with a sale indicator
sf_parcels <- chars_data %>%
  filter(!meta_class %in% c("211", "212")) %>%
  # Join on sales indicator for most recent sale in the last two years -
  # if there is one
  left_join(
    training_data %>%
      filter(
        year >= as.numeric(params$year) - 1,
        !sv_is_outlier,
        meta_sale_date == max(meta_sale_date, na.rm = TRUE),
        .by = meta_pin
      ) %>%
      distinct() %>% # we shouldn't need this here, why are there duplicate
      # sales in the training data?
      select(meta_pin, meta_sale_date, meta_year)
  ) %>%
  mutate(
    sale = factor(
      as.numeric(!is.na(meta_sale_date)),
      levels = c(0, 1),
      labels = c("No Sale", "Sale")
    ),
    char_recent_renovation = case_when(
      char_recent_renovation == 0 ~ "No",
      TRUE ~ "Yes"
    ),
    across(
      (starts_with("char_") &
        where(is.character)) | c(
        char_recent_renovation,
        char_type_resd
      ),
      as.factor
    )
  ) %>%
  select(sale, all_of(vars)) %>%
  vars_recode(
    type = "long",
    dict = ccao::vars_dict_legacy
  ) %>%
  vars_rename(names_from = "athena", names_to = "pretty") %>%
  rename_with(~ gsub("loc_", "", .x)) %>%
  rename_with(~ gsub("_", " ", .x)) %>%
  rename_with(toTitleCase) %>%
  rename_with(~ gsub("Acs5", "ACS5", .x))

logit_parcels <- sf_parcels %>%
  mutate(across(where(is.factor), ~ addNA(.x))) %>%
  drop_na()

vars <- names(sf_parcels) %>%
  grep("Sale|Township", ., value = TRUE, invert = TRUE)
```


## Logit Regressions

All covariates regressed on sale status - significance indicates a non-random effect on the likelihood of a sale ceterus paribus. Sample is universe of single family parcels from the past two years, excluding some parcels with incomplete characteristics.

```{r logit}
remove <-
  names(sf_parcels %>% rename_with(~ gsub(" |-", "_", .x)) %>%
    select(where(is.factor))) %>%
  grep(
    "Sale|Township|Apartments|Single|Commercial",
    .,
    invert = TRUE,
    value = TRUE
  )

fmla <- as.formula(
  paste0("Sale ~ ", paste(gsub(" |-", "_", vars), collapse = "+"))
  )

logit <- glm(fmla,
  data = logit_parcels %>%
    rename_with(~ gsub(" |-", "_", .x)),
  family = "binomial"
)

sample_size <- logit$df.null

county_logit <- logit %>%
  broom::tidy() %>%
  mutate(
    p.value = paste0(scales::pvalue(p.value), (stars.pval(p.value))),
    term = gsub("_", " ", gsub(paste(remove, collapse = "|"), "", term))
  ) %>%
  kable(
    caption = paste0(
      "Sample size: ",
      format(sample_size, big.mark = ","),
      " single-family parcels"
    ),
    col.names = c("Predictor", "β", "SE", "t", "p"),
    digits = c(0, 5, 5, 3, 0),
    align = c("l", "r", "c", "r", "l"),
    padding = 5L
  ) %>%
  add_footnote(
    "p-value notation: 0-0.001 = ***, 0.001-0.01 = **, 0.01-0.05 = *, 0.05-0.1 = .", #noqa
    notation = "none"
  ) %>%
  kable_styling("striped",
    fixed_thead = TRUE,
    html_font = "Consolas"
  ) %>%
  pack_rows("Type of Residence", 12, 20) %>%
  pack_rows("Cathedral Ceiling", 21, 22) %>%
  pack_rows("Attic Finish", 23, 25) %>%
  pack_rows("Garage 1 Attached", 26, 27) %>%
  pack_rows("Garage 1 Area Included", 28, 29) %>%
  pack_rows("Garage 1 Size", 30, 37) %>%
  pack_rows("Garage 1 Ext. Wall Material", 38, 41) %>%
  pack_rows("Attic Type", 42, 44) %>%
  pack_rows("Basement Type", 45, 48) %>%
  pack_rows("Exterior Wall Material", 49, 51) %>%
  pack_rows("Central Heating", 52, 55) %>%
  pack_rows("Basement Finish", 56, 58) %>%
  pack_rows("Roof Material", 59, 64) %>%
  pack_rows("Porch", 65, 66) %>%
  pack_rows("Central Air Conditioning", 67, 68) %>%
  pack_rows("Design Plan", 69, 70) %>%
  pack_rows("Recent Renovation", 71, 71)

logits <- logit_parcels %>%
  split(.$`Meta Township Name`) %>%
  sapply(function(x) {
    if (nrow(x) < 1000) {
      trim_vars <- grep("Renovation", vars, value = TRUE, invert = TRUE)

      fmla <- as.formula(paste0("Sale ~ ", paste0(gsub(
        " |-", "_", trim_vars
      ), collapse = "+")))

      glm(fmla,
        data = x %>%
          rename_with(~ gsub(" |-", "_", .x)),
        family = "binomial"
      ) %>%
        broom::tidy() %>%
        mutate(
          p.value = paste0(scales::pvalue(p.value), (stars.pval(p.value))),
          term = gsub("_", " ", term)
        ) %>%
        kable(
          col.names = c("Predictor", "β", "SE", "t", "p"),
          digits = c(0, 5, 5, 3, 0),
          align = c("l", "r", "c", "r", "l"),
          padding = 5L
        ) %>%
        add_footnote(
          "p-value notation: 0-0.001 = ***, 0.001-0.01 = **, 0.01-0.05 = *, 0.05-0.1 = .", # noqa
          notation = "none"
        ) %>%
        kable_styling(
          "striped",
          position = "left",
          fixed_thead = TRUE,
          html_font = "Consolas"
        )
    } else {
      fmla <- as.formula(
        paste0("Sale ~ ", paste0(gsub(" |-", "_", vars), collapse = "+"))
        )

      glm(fmla,
        data = x %>%
          rename_with(~ gsub(" |-", "_", .x)),
        family = "binomial"
      ) %>%
        broom::tidy() %>%
        mutate(
          p.value = paste0(scales::pvalue(p.value), (stars.pval(p.value))),
          term = gsub("_", " ", term)
        ) %>%
        kable(
          col.names = c("Predictor", "β", "SE", "t", "p"),
          digits = c(0, 5, 5, 3, 0),
          align = c("l", "r", "c", "r", "l"),
          padding = 5L
        ) %>%
        add_footnote(
          "p-value notation: 0-0.001 = ***, 0.001-0.01 = **, 0.01-0.05 = *, 0.05-0.1 = .", #noqa
          notation = "none"
        ) %>%
        kable_styling(
          "striped",
          position = "left",
          fixed_thead = TRUE,
          html_font = "Consolas"
        )
    }
  }, simplify = FALSE)
```

::: panel-tabset
## County

```{r}
county_logit
```

## Barrington

```{r}
logits[["Barrington"]]
```

## Berwyn

```{r}
logits[["Berwyn"]]
```

## Bloom

```{r}
logits[["Bloom"]]
```

## Bremen

```{r}
logits[["Bremen"]]
```

## Calumet

```{r}
logits[["Calumet"]]
```

## Cicero

```{r}
logits[["Cicero"]]
```

## Elk Grove

```{r}
logits[["Elk Grove"]]
```

## Evanston

```{r}
logits[["Evanston"]]
```

## Hanover

```{r}
logits[["Hanover"]]
```

## Hyde Park

```{r}
logits[["Hyde Park"]]
```

## Jefferson

```{r}
logits[["Jefferson"]]
```

## Lake

```{r}
logits[["Lake"]]
```

## Lake View

```{r}
logits[["Lake View"]]
```

## Lemont

```{r}
logits[["Lemont"]]
```

## Leyden

```{r}
logits[["Leyden"]]
```

## Lyons

```{r}
logits[["Lyons"]]
```

## Maine

```{r}
logits[["Maine"]]
```

## New Trier

```{r}
logits[["New Trier"]]
```

## Niles

```{r}
logits[["Niles"]]
```

## North Chicago

```{r}
logits[["North Chicago"]]
```

## Northfield

```{r}
logits[["Northfield"]]
```

## Norwood Park

```{r}
logits[["Norwood Park"]]
```

## Oak Park

```{r}
logits[["Oak Park"]]
```

## Orland

```{r}
logits[["Orland"]]
```

## Palatine

```{r}
logits[["Palatine"]]
```

## Palos

```{r}
logits[["Palos"]]
```

## Proviso

```{r}
logits[["Proviso"]]
```

## Rich

```{r}
logits[["Rich"]]
```

## River Forest

```{r}
logits[["River Forest"]]
```

## Riverside

```{r}
logits[["Riverside"]]
```

## Rogers Park

```{r}
logits[["Rogers Park"]]
```

## Schaumburg

```{r}
logits[["Schaumburg"]]
```

## South Chicago

```{r}
logits[["South Chicago"]]
```

## Stickney

```{r}
logits[["Stickney"]]
```

## Thornton

```{r}
logits[["Thornton"]]
```

## West Chicago

```{r}
logits[["West Chicago"]]
```

## Wheeling

```{r}
logits[["Wheeling"]]
```

## Worth

```{r}
logits[["Worth"]]
```
:::

## Standardized mean differences

The standardized (mean) difference is a measure of distance between two group
means in terms of one or more variables.

$$SMD = {\text{Difference in mean outcome between groups} \over \text{Standard
deviation of outcome among participants}}$$

We're looking for SMDs greater than .1 or .05 for sensitive covariates.

```{r SMDs}
smd_table <- CreateTableOne(
  vars = vars %>% grep("Median", ., invert = TRUE, value = TRUE),
  strata = "Sale",
  data = sf_parcels,
  test = FALSE,
  includeNA = TRUE
)

print(smd_table, smd = TRUE)
```

## Empirical Cumulative Distributions

For the empirical cumulative distributions, we are looking for overlap/similar
distributions. Numeric variables scaled and centered.

::: panel-tabset
## Numeric columns:

```{r ECDFs1}
grobs <-
  lapply(names(sf_parcels %>% select(where(is.numeric))), function(x) {
    data <- sf_parcels %>% select(Sale, any_of(x))
    ggplot(data, aes(scale(sf_parcels %>% pull(x)),
      color = Sale
    )) +
      stat_ecdf(geom = "step") +
      labs(x = str_wrap(x, width = 25)) +
      theme_minimal() +
      theme(
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none"
      )
  })

for (i in seq(1, length(grobs), by = 6)) {
  if (i == 13) {
    grid.arrange(
      grobs = list(
        grobs[[i]],
        grobs[[i + 1]],
        grobs[[i + 2]],
        grobs[[i + 3]],
        rectGrob(gp = gpar(col = "white")),
        rectGrob(gp = gpar(col = "white"))
      ),
      ncol = 3,
      nrow = 2
    )
  } else {
    grid.arrange(
      grobs = list(
        grobs[[i]], grobs[[i + 1]], grobs[[i + 2]],
        grobs[[i + 3]], grobs[[i + 4]], grobs[[i + 5]]
      ),
      ncol = 3,
      nrow = 2
    )
  }
}
```

## Factor columns:

```{r ECDFs2}
grobs <-
  lapply(names(sf_parcels %>% select(where(is.factor) &
    any_of(vars))), function(x) {
    ggplot(sf_parcels, aes(as.numeric(sf_parcels %>% pull(x)),
      color = Sale
    )) +
      stat_ecdf(geom = "step") +
      labs(x = str_wrap(x, width = 25)) +
      theme_minimal() +
      theme(
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none"
      )
  })

for (i in seq(1, length(grobs), by = 6)) {
  if (i == 13) {
    grid.arrange(
      grobs = list(
        grobs[[i]], grobs[[i + 1]], grobs[[i + 2]],
        grobs[[i + 3]], grobs[[i + 4]],
        rectGrob(gp = gpar(col = "white"))
      ),
      ncol = 3,
      nrow = 2
    )
  } else {
    grid.arrange(
      grobs = list(
        grobs[[i]], grobs[[i + 1]], grobs[[i + 2]],
        grobs[[i + 3]], grobs[[i + 4]], grobs[[i + 5]]
      ),
      ncol = 3,
      nrow = 2
    )
  }
}
```

:::
