```{r}
create_leaflet_map <- function(dataset, legend_value, legend_title, order_scheme = "high", longitude = "loc_longitude", latitude = "loc_latitude") {
  # Filter neighborhoods that have at least one observation
  nbhd_borders <- nbhd %>%
    right_join(dataset, by = c("town_nbhd" = "meta_nbhd_code"))

  # Create the color palette based on order_scheme
  if (order_scheme == "low") {
    pal <- colorNumeric(palette = "Reds", domain = dataset[[legend_value]], reverse = TRUE)
  } else {
    pal <- colorNumeric(palette = "Reds", domain = dataset[[legend_value]])
  }

  # Calculate the bounding box of the filtered neighborhoods
  bbox <- st_bbox(nbhd_borders)

  # Create the leaflet map
  leaflet(dataset) %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addCircleMarkers(
      lng = ~ get(longitude),
      lat = ~ get(latitude),
      radius = 5,
      color = ~ pal(dataset[[legend_value]]),
      popup = ~ {
        shap_values <- dataset %>%
          select(starts_with("target_feature_shap_")) %>%
          summarise_all(~ ifelse(!is.na(.), sprintf("SHAP: %.2f", .), NA)) %>%
          apply(1, function(row) {
            paste(na.omit(row), collapse = "<br>")
          })

        paste(
          "Pin: ", meta_pin,
          ifelse(shap_values == "", "", paste0("<br>", shap_values)),
          "<br>", "Relative SHAP: ", scales::percent(relative_shap, accuracy = 0.01),
          "<br>", "Feature: ", sprintf("%.2f", get(params$added_feature)),
          "<br>", "New FMV: ", pred_pin_final_fmv_new,
          "<br>", "Comparison FMV: ", pred_pin_final_fmv_comp,
          "<br>", "FMV Difference: ", scales::percent(diff_pred_pin_final_fmv)
        )
      }
    ) %>%
    addPolygons(
      data = nbhd_borders,
      color = "black",
      weight = 2,
      fill = FALSE
    ) %>%
    addLegend(
      "bottomright",
      pal = pal,
      values = dataset[[legend_value]],
      title = legend_title
    )
}




```

## Highest and Lowest 100 Values

Three different types of high and low values are produced; the values of the feature we are analyzing, the impact that can be determined through the SHAPs, and the largest effects in change in FMV.

::: panel-tabset
### Largest 100 Values

Be careful interpreting values which are the max and min of the raw value, since ties are not accounted for. For example, if there are 10,000 parcels which are 0 feet from a newly constructed building, the map will not be a full representation.

```{r}
highest_100 <- leaflet_data %>%
   arrange(desc(!!sym(target_feature_value))) %>%
   dplyr::slice(1:100)

create_leaflet_map(highest_100, {{target_feature_value}}, "Largest 100 Values")
```

### Lowest 100 Values

Be careful interpreting values which are the max and min of the raw value, since ties are not accounted for. For example, if there are 10,000 parcels which are 0 feet from a newly constructed building, the map will not be a full representation.

```{r}
lowest_100 <- leaflet_data %>%
  distinct(meta_pin, .keep_all = TRUE) %>%
  arrange(!!sym({{target_feature_value}})) %>%
  slice(1:100)

create_leaflet_map(lowest_100, {{target_feature_value}}, "Lowest 100 Values", order_scheme = "low")
```

### Highest 100 SHAP Values

```{r}
highest_100 <- leaflet_data %>%
  arrange(desc(shap_total)) %>%
  slice(1:100)

create_leaflet_map(highest_100, "shap_total", "Highest 100 SHAPs")
```

### Lowest 100 SHAP Values

```{r}
lowest_100 <- leaflet_data %>%
  arrange(shap_total) %>%
  slice(1:100)

create_leaflet_map(lowest_100, "shap_total", "Lowest 100 SHAPs", order_scheme = "low")
```
:::

## Largest FMV Changes

Multicard parcels have heuristic which limits their change. The added feature may trigger (or not trigger it), leading to changes much larger than the added feature's impact.

::: panel-tabset
### 100 Largest FMV Increases

```{r}
largest_fmv_increases <- leaflet_data %>%
  arrange(desc(diff_pred_pin_final_fmv)) %>%
  slice(1:100)

# Call the function with the pre-sliced dataset
create_leaflet_map(largest_fmv_increases, "diff_pred_pin_final_fmv", "Largest FMV Increases (%)")
```

### 100 Largest FMV Decreases

Multicard parcels have heuristic which limits their change. The added feature may trigger (or not trigger it), leading to changes much larger than the added feature's impact.

```{r}
largest_fmv_decreases <- leaflet_data %>%
  arrange(diff_pred_pin_final_fmv) %>%
  slice(1:100)

create_leaflet_map(largest_fmv_decreases, "diff_pred_pin_final_fmv", "Largest FMV Decreases (%)", order_scheme = "low")
```

### 100 Largest FMV Initial Increases

```{r}
largest_fmv_increases <- leaflet_data %>%
  arrange(desc(diff_pred_pin_initial_fmv)) %>%
  slice(1:100)

# Call the function with the pre-sliced dataset
create_leaflet_map(largest_fmv_increases, "diff_pred_pin_initial_fmv", "Largest FMV Increases (%)")
```

### 100 Largest Initial FMV Decreases

```{r}
largest_fmv_decreases <- leaflet_data %>%
  arrange(diff_pred_pin_initial_fmv) %>%
  slice(1:100)

create_leaflet_map(largest_fmv_decreases, "diff_pred_pin_initial_fmv", "Largest FMV Decreases (%)", order_scheme = "low")
```

## Largest FMV Increases no Multicards

```{r}
largest_fmv_increases <- leaflet_data %>%
  group_by(meta_pin) %>%
  filter(n() == 1) %>%
  ungroup() %>%
  arrange(desc(diff_pred_pin_final_fmv)) %>%
  slice(1:100)

create_leaflet_map(largest_fmv_increases, "diff_pred_pin_final_fmv", "Largest FMV Increases")
```

## Largest FMV Decreases no Multicards

```{r}
largest_fmv_decreases <- leaflet_data %>%
  group_by(meta_pin) %>%
  filter(n() == 1) %>%
  ungroup() %>%
  arrange(diff_pred_pin_initial_fmv) %>%
  slice(1:100)

create_leaflet_map(largest_fmv_increases, "diff_pred_pin_final_fmv", "Largest FMV Increases (%)", order_scheme = "low")
```
:::

## Neighborhoods with the Highest and Lowest SHAP Values

These maps identify neighborhoods where the added feature is having the largest impact on SHAP values. By selecting neighborhoods with the highest mean(absolute value), you can take a closer look at how individual parcels in these neighborhoods are affected.

::: panel-tabset
```{r processing_SHAP_values}
selected_data <- leaflet_data %>%
  group_by(meta_nbhd_code) %>%
  mutate(mean_value = mean(abs(shap_total)), na.rm = TRUE) %>%
  ungroup() %>%
  distinct(meta_nbhd_code, .keep_all = TRUE) %>%
  arrange(mean_value)

# Select top 2 and bottom 2 neighborhoods based on mean SHAP values
selected_nbhd_codes <- selected_data %>%
  slice(c(1:2, (n() - 1):n())) %>%
  pull(meta_nbhd_code)


filtered_data <- filter(leaflet_data, meta_nbhd_code %in% selected_nbhd_codes)


# Separate high and low mean value neighborhoods
high_mean_data <- filtered_data %>%
  filter(meta_nbhd_code %in% selected_nbhd_codes[(length(selected_nbhd_codes) - 1):length(selected_nbhd_codes)])

low_mean_data <- filtered_data %>%
  filter(meta_nbhd_code %in% selected_nbhd_codes[1:2])
```

### 2 Highest SHAP Neighborhoods

```{r}
create_leaflet_map(high_mean_data, "shap_total", "SHAP Values")
```

### 2 Lowest SHAP Neighborhoods

```{r}
create_leaflet_map(low_mean_data, "shap_total", "SHAP Values")
```
:::
