
:::

## Correlation Between Added Feature and Other Features

Here, the goal is to see if the added feature *very* neatly aligns with other existing features. Columns are produced with both the absolute value of the correlation (for easy sorting), as well as the correlation to help decipher the direction of the relationship.

```{r correlation between features}
clean_column_values <- function(df, column_name) {
  df[[column_name]] <- df[[column_name]] %>%
    gsub("^meta_|^prox_|^other_|^loc_|^char_|^acs5|^acs_|^ccao_", "", .) %>%
    gsub("_", " ", .) %>%
    stringr::str_to_title()
  return(df)
}

columns_to_remove <- c(
  "Sale Year",
  "Sale Month of Year",
  "Sale Day of Year",
  "Sale Day of Week",
  "Sale Day of Month")

if (params$type == "continuous") {
    
  numeric_cols <- assessment_data_new %>%
    rename('Property Class Meta' = meta_class,
           'Property Class Char' = char_class) %>%
    ccao::vars_rename(names_from = "model",
                      names_to = "pretty",
                      type = "inplace",
                      dict = ccao::vars_dict) %>%
    select_if(is.numeric) %>%
    select(-all_of(columns_to_remove))

  # Initialize a data frame to store correlation results
  correlation_results <- data.frame(Feature = character(), Correlation = numeric(), Abs_Correlation = numeric(), stringsAsFactors = FALSE)

  # Loop through each numeric column and calculate correlation and absolute correlation
  for (col_name in names(numeric_cols)) {
    # Filter out rows with missing values in the two columns being compared
    complete_cases <- complete.cases(numeric_cols[[col_name]], assessment_data_new[[params$added_feature]])

    # Only compute correlation if there are complete cases
    if (sum(complete_cases) > 0) {
      correlation_value <- cor(numeric_cols[[col_name]][complete_cases], assessment_data_new[[params$added_feature]][complete_cases], use = "complete.obs")
      abs_correlation_value <- abs(cor(abs(numeric_cols[[col_name]][complete_cases]), abs(assessment_data_new[[params$added_feature]][complete_cases]), use = "complete.obs"))
      correlation_results <- rbind(correlation_results, data.frame(Feature = col_name, Correlation = correlation_value, Abs_Correlation = abs_correlation_value))
    }
  }

  # Sort the correlation results in descending order by Correlation
  correlation_results <- correlation_results %>%
    arrange(dplyr::desc(Abs_Correlation)) %>%
    mutate(across(where(is.numeric), ~ round(., 2)))

  top_10_features <- correlation_results %>%
    slice(1:10) %>%
    pull(Feature)
  
  correlation_results_clean <- correlation_results %>%
  clean_column_values("Feature") %>%
  slice(2:n())
  

  # Display the correlation results as a scrollable table
  datatable(correlation_results_clean, options = list(scrollX = TRUE, scrollY = TRUE, pageLength = 10, order = list(list(1, "desc"))))
} else {
  print(paste("assessment_data_new$", params$added_feature, " is not numeric.", sep = ""))
}
```

## Correlation Plot

This selects the 10 most correlated features (in terms of absolute value) from the previous chart and creates a correlation plot
```{r}
# Select the top 10 features, remove rows with NA values, rename columns, calculate the correlation, and plot the correlation matrix
assessment_data_new %>%
      rename('Property Class Meta' = meta_class,
           'Property Class Char' = char_class) %>%
    ccao::vars_rename(names_from = "model",
                      names_to = "pretty",
                      type = "inplace",
                      dict = ccao::vars_dict) %>%
    select_if(is.numeric) %>%
    select(-all_of(columns_to_remove)) %>%
  select(all_of(top_10_features)) %>%
  na.omit() %>%
  rename_with(~ gsub("^meta_|^prox_|^other_|^loc_|^char_|^acs5|^acs_|^ccao_", "", .)) %>%
  rename_with(~ gsub("_", " ", .)) %>%
  cor() %>%
  corrplot(method = "circle", tl.cex = 0.6, tl.srt = 45, addgrid.col = "grey", mar = c(1, 1, 1, 1))
```
