
# Descriptive Statistics

::: {.panel-tabset}

```{r}
# Function to create summary tables
create_summary_table <- function(data, target_feature, group_by_column = NULL) {
  target_feature <- sym(target_feature)

  summary_data <- if (!is.null(group_by_column)) {
    data %>%
      group_by(!!sym(group_by_column)) %>%
      summarize(
        Mean = round(mean(!!target_feature, na.rm = TRUE), 2),
        Median = round(median(!!target_feature, na.rm = TRUE), 2),
        `10th Percentile` = round(quantile(!!target_feature, 0.1, na.rm = TRUE), 2),
        `90th Percentile` = round(quantile(!!target_feature, 0.9, na.rm = TRUE), 2),
        Mode = round(as.numeric(names(sort(table(!!target_feature), decreasing = TRUE)[1])), 2)
      )
  } else {
    data %>%
      summarize(
        Mean = round(mean(!!target_feature, na.rm = TRUE), 2),
        Median = round(median(!!target_feature, na.rm = TRUE), 2),
        `10th Percentile` = round(quantile(!!target_feature, 0.1, na.rm = TRUE), 2),
        `90th Percentile` = round(quantile(!!target_feature, 0.9, na.rm = TRUE), 2),
        Mode = round(as.numeric(names(sort(table(!!target_feature), decreasing = TRUE)[1])), 2)
      )
  }

  # Display the summary in a datatable
  datatable(
    summary_data,
    options = list(
      scrollY = "300px",
      scrollX = TRUE,
      paging = FALSE,
      searching = TRUE
    ),
    rownames = FALSE
  )
}
```

## Overall

```{r mean_median}
# For Overall
create_summary_table(pin_individual, target_feature = {{ target_feature_value }})
```

## Township
```{r mean_median_township}
# For Township
create_summary_table(pin_individual, target_feature = {{ target_feature_value }}, group_by_column = "meta_township_name")
```

## Neighborhood

```{r mean_median_neighborhood}
# For Neighborhood
create_summary_table(pin_individual, target_feature = {{ target_feature_value }}, group_by_column = "meta_nbhd_code")
```

:::

## Histogram

```{r }
pin_individual %>%
  mutate(
    mean_value = mean(!!sym({{ target_feature_value }}), na.rm = TRUE),
    median_value = median(!!sym({{ target_feature_value }}), na.rm = TRUE)
  ) %>%
  ggplot(aes(x = !!sym({{ target_feature_value }}))) +
  geom_histogram(fill = "blue", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = mean_value, color = "Mean"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  geom_vline(aes(xintercept = median_value, color = "Median"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  scale_color_manual(
    name = "Statistics",
    values = c(Mean = "red", Median = "green"),
    labels = c("Mean", "Median")
  ) +
  labs(
    x = {{ target_feature_value }},
    y = "Frequency"
  ) +
  theme_minimal()
```


## FMV Change Histogram

This chart shows the distribution of the value of 'diff_pred_pin_initial_fmv' in the model with the added feature minus the model without the added feature. Outliers outside of 95% are removed to make the chart more readable. The largest 100 increases and decreases are displayed in maps in section X. 
```{r}
pin_individual %>%
  filter(
    diff_pred_pin_initial_fmv >= quantile(diff_pred_pin_initial_fmv, 0.025, na.rm = TRUE) &
      diff_pred_pin_initial_fmv <= quantile(diff_pred_pin_initial_fmv, 0.975, na.rm = TRUE)
  ) %>%
  mutate(
    mean_value = mean(diff_pred_pin_initial_fmv, na.rm = TRUE),
    median_value = median(diff_pred_pin_initial_fmv, na.rm = TRUE)
  ) %>%
  ggplot(aes(x = diff_pred_pin_initial_fmv)) +
  geom_histogram(fill = "blue", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = mean_value, color = "Mean"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  geom_vline(aes(xintercept = median_value, color = "Median"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  scale_color_manual(
    name = "Statistics",
    values = c(Mean = "red", Median = "green"),
    labels = c("Mean", "Median")
  ) +
  labs(
    x = "Change in FMV",
    y = "Frequency"
  ) +
  theme_minimal()
```

## SHAP Histogram

```{r}
shap_new %>%
  filter(
    !!sym({{ target_feature_value }}) >= quantile(!!sym({{ target_feature_value }}), 0.025, na.rm = TRUE) &
      !!sym({{ target_feature_value }}) <= quantile(!!sym({{ target_feature_value }}), 0.975, na.rm = TRUE)
  ) %>%
  mutate(
    mean_value = mean(!!sym({{ target_feature_value }}), na.rm = TRUE),
    median_value = median(!!sym({{ target_feature_value }}), na.rm = TRUE)
  ) %>%
  ggplot(aes(x = !!sym({{ target_feature_value }}))) +
  geom_histogram(fill = "blue", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = mean_value, color = "Mean"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  geom_vline(aes(xintercept = median_value, color = "Median"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  scale_color_manual(
    name = "Statistics",
    values = c(Mean = "red", Median = "green"),
    labels = c("Mean", "Median")
  ) +
  labs(
    x = "SHAP Value",
    y = "Frequency"
  ) +
  theme_minimal()
```

:::
