
# Descriptive Statistics


::: {.panel-tabset}


## Overall

```{r mean_median}
pin_individual %>%
    summarize(
    Mean = round(mean(!!sym({{target_feature_value}}), na.rm = TRUE), 2),
    Median = round(median(!!sym({{target_feature_value}}), na.rm = TRUE), 2),
    '10th Percentile' = round(quantile(!!sym({{target_feature_value}}), 0.1, na.rm = TRUE), 2),
    '90th Percentile' = round(quantile(!!sym({{target_feature_value}}), 0.9, na.rm = TRUE), 2),
    Mode = round(as.numeric(names(sort(table(!!sym({{target_feature_value}})), decreasing = TRUE)[1])), 2)) %>%
    datatable(
    options = list(
      scrollY = "300px",
      scrollX = TRUE,
      paging = FALSE,
      searching = TRUE
    ),
    rownames = FALSE
  )
```

## Township
```{r mean_median_township}
pin_individual %>%
    group_by(meta_township_name) %>%
    summarize(
    Mean = round(mean(!!sym({{target_feature_value}}), na.rm = TRUE), 2),
    Median = round(median(!!sym({{target_feature_value}}), na.rm = TRUE), 2),
    '10th Percentile' = round(quantile(!!sym({{target_feature_value}}), 0.1, na.rm = TRUE), 2),
    '90th Percentile' = round(quantile(!!sym({{target_feature_value}}), 0.9, na.rm = TRUE), 2),
    Mode = round(as.numeric(names(sort(table(!!sym({{target_feature_value}})), decreasing = TRUE)[1])), 2)) %>%
    datatable(
    options = list(
      scrollY = "300px",
      scrollX = TRUE,
      paging = FALSE,
      searching = TRUE
    ),
    rownames = FALSE
  )
```


## Neighborhood

```{r mean_median_neighborhood}
pin_individual %>%
    group_by(meta_nbhd_code) %>%
    summarize(
    Mean = round(mean(!!sym({{target_feature_value}}), na.rm = TRUE), 2),
    Median = round(median(!!sym({{target_feature_value}}), na.rm = TRUE), 2),
    '10th Percentile' = round(quantile(!!sym({{target_feature_value}}), 0.1, na.rm = TRUE), 2),
    '90th Percentile' = round(quantile(!!sym({{target_feature_value}}), 0.9, na.rm = TRUE), 2),
    Mode = round(as.numeric(names(sort(table(!!sym({{target_feature_value}})), decreasing = TRUE)[1])), 2)) %>%
    datatable(
    options = list(
      scrollY = "300px",
      scrollX = TRUE,
      paging = FALSE,
      searching = TRUE
    ),
    rownames = FALSE
  )
```

:::

## Histogram

```{r }
pin_individual %>%
    mutate(
      mean_value = mean(!!sym({{target_feature_value}}), na.rm = TRUE),
      median_value = median(!!sym({{target_feature_value}}), na.rm = TRUE)
    ) %>%
    ggplot(aes(x = !!sym({{target_feature_value}}))) +
    geom_histogram(fill = "blue", color = "black", alpha = 0.7) +
    geom_vline(aes(xintercept = mean_value, color = "Mean"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
    geom_vline(aes(xintercept = median_value, color = "Median"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
    scale_color_manual(
      name = "Statistics",
      values = c(Mean = "red", Median = "green"),
      labels = c("Mean", "Median")
    ) +
    labs(
      x = {{target_feature_value}},
      y = "Frequency"
    ) +
    theme_minimal()

```


## FMV Change Histogram

This chart shows the distribution of the value of 'diff_pred_pin_initial_fmv' in the model with the added feature minus the model without the added feature. Outliers outside of 95% are removed to make the chart more readable. The largest 100 increases and decreases are displayed in maps in section X. 
```{r}
pin_individual %>%
  filter(
    diff_pred_pin_initial_fmv >= quantile(diff_pred_pin_initial_fmv, 0.025, na.rm = TRUE) &
      diff_pred_pin_initial_fmv <= quantile(diff_pred_pin_initial_fmv, 0.975, na.rm = TRUE)
  ) %>%
  mutate(
    mean_value = mean(diff_pred_pin_initial_fmv, na.rm = TRUE),
    median_value = median(diff_pred_pin_initial_fmv, na.rm = TRUE)
  ) %>%
  ggplot(aes(x = diff_pred_pin_initial_fmv)) +
  geom_histogram(fill = "blue", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = mean_value, color = "Mean"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  geom_vline(aes(xintercept = median_value, color = "Median"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  scale_color_manual(
    name = "Statistics",
    values = c(Mean = "red", Median = "green"),
    labels = c("Mean", "Median")
  ) +
  labs(
    x = "Change in FMV",
    y = "Frequency"
  ) +
  theme_minimal()
```

## SHAP Histogram

```{r}
shap_new %>%
  filter(
    !!sym({{target_feature_value}}) >= quantile(!!sym({{target_feature_value}}), 0.025, na.rm = TRUE) &
      !!sym({{target_feature_value}}) <= quantile(!!sym({{target_feature_value}}), 0.975, na.rm = TRUE)
  ) %>%
  mutate(
    mean_value = mean(!!sym({{target_feature_value}}), na.rm = TRUE),
    median_value = median(!!sym({{target_feature_value}}), na.rm = TRUE)
  ) %>%
  ggplot(aes(x = !!sym({{target_feature_value}}))) +
  geom_histogram(fill = "blue", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = mean_value, color = "Mean"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  geom_vline(aes(xintercept = median_value, color = "Median"), linetype = "dashed", linewidth = 1, show.legend = TRUE) +
  scale_color_manual(
    name = "Statistics",
    values = c(Mean = "red", Median = "green"),
    labels = c("Mean", "Median")
  ) +
  labs(
    x = "SHAP Value",
    y = "Frequency"
  ) +
  theme_minimal()
```

:::
