---
title: "Balance Testing Residential Parcels Across Sale Status"
execute:
  echo: false
  warning: false
format:
  html:
    embed-resources: true
    toc: true
    toc_float: true
    fig-align: center
    fontsize: 12pt
editor: source
---

```{r setup}

# This script runs balance tests comparing the universe of residential parcels
# to the sample of residential parcels with sales in the previous 8 years
# that are used to train the data department's CAMA.

# Load necessary libraries
library(arrow)
library(ccao)
library(DBI)
library(dplyr)
library(ggplot2)
library(glue)
library(grid)
library(gridExtra)
library(gtools)
library(kableExtra)
library(noctua)
library(plotly)
library(purrr)
library(scales)
library(sf)
library(stringr)
library(tableone)
library(tidyr)
library(tools)

options(width = 120)
options(knitr.kable.NA = '')

# Connect to Athena
AWS_ATHENA_CONN_NOCTUA <- dbConnect(noctua::athena())

# Declare time frame for study
time_frame <- c("min" = "2021", "max" = "2022")

# Data path
data_path <- "~/data/res_parcels.parquet"

# GATHER DATA ----

sales <- arrow::read_parquet("~/ccao_res_avm/input/training_data.parquet")

# All residential parcel characteristics from time frame, with a sale indicator

if (!file.exists(data_path)) {
  
  dbGetQuery(
  conn = AWS_ATHENA_CONN_NOCTUA,
  glue("

    SELECT * FROM model.vw_card_res_input vcri
    WHERE vcri.meta_year BETWEEN '{time_frame['min']}' AND '{time_frame['max']}'
    AND meta_class NOT IN ('211', '212')

  ")
) %>%
  write_parquet(data_path)
  
}

sf_parcels <- read_parquet(data_path) %>%
  # Join on sales indicator for most recent sale in the last two years -
  # if there is one
  left_join(
    sales %>%
      filter(
        dplyr::between(meta_year, time_frame['min'], time_frame['max']) &
          sv_is_outlier == FALSE
        ) %>%
      select(meta_pin, meta_sale_date, meta_year) %>%
      group_by(meta_pin) %>%
      filter(meta_sale_date == max(meta_sale_date, na.rm = TRUE)) %>%
      ungroup()
  ) %>%
  mutate(sale = as.numeric(!is.na(meta_sale_date)))
  

# Vars to test across
vars <- grep("char_|3435|acs5_median|township_name", names(sf_parcels), value = TRUE) %>%
  grep("percentile|_apts|_ncu|_use|_qlty", ., value = TRUE, invert = TRUE)

sf_parcels <- sf_parcels %>%
  mutate(
    char_recent_renovation = case_when(
      char_recent_renovation == 0 ~ "No",
      TRUE ~ "Yes"
      )
  ) %>%
  mutate(
    sale = factor(sale, levels = c(0, 1), labels = c("No Sale", "Sale")),
    across(starts_with("char_") & where(is.character), ~ as.factor(.x)),
    across(c(
      loc_cook_municipality_name,
      loc_ward_num,
      char_recent_renovation,
      char_type_resd
    ), as.factor)
  ) %>%
  select(sale, all_of(vars)) %>%
  vars_recode(
    type = "long",
    dict = ccao::vars_dict_legacy
  ) %>%
  vars_rename(names_from = 'athena', names_to = 'pretty') %>%
  rename_with(~ gsub("loc_", "", .x)) %>%
  rename_with(~ gsub("_", " ", .x)) %>%
  rename_with(toTitleCase) %>%
  rename_with(~ gsub("Acs5", "ACS5", .x))

logit_parcels <- sf_parcels %>%
  mutate(across(where(is.factor), ~ addNA(.x))) %>%
  drop_na()

vars <- names(sf_parcels) %>%
  grep("Sale|Township", ., value = TRUE, invert = TRUE)

```

## Logit Regressions

All covariates regressed on sale status - significance indicates a non-random effect on the likelihood of a sale ceterus paribus.  Sample is universe of '21 & '22 single family parcels, excluding some parcels with incomplete characteristics.

```{r logit}

remove <- names(sf_parcels %>% rename_with(~ gsub(" |-", "_", .x)) %>% select(where(is.factor))) %>%
  grep("Sale|Township|Apartments|Single|Commercial", ., invert = TRUE, value = TRUE)

fmla <- as.formula(
        paste0("Sale ~ ", paste(gsub(" |-", "_", vars), collapse = "+"))
        )

logit <- glm(
      fmla,
      data = logit_parcels %>%
        rename_with(~ gsub(" |-", "_", .x)), family = "binomial"
      )

sample_size <- logit$df.null

county_logit <- logit %>%
  broom::tidy() %>%

  mutate(
    p.value = paste0(scales::pvalue(p.value), (stars.pval(p.value))),
    term = gsub("_", " ", gsub(paste(remove, collapse = "|"), "", term))
  ) %>%
  kable(
    caption = paste0("Sample size: ", format(sample_size, big.mark = ","), " single-family parcels"),
    col.names = c("Predictor", "β", "SE", "t", "p"),
    digits = c(0, 5, 5, 3, 0),
    align = c("l", "r", "c", "r", "l"),
    padding = 5L
  ) %>%
  add_footnote(
    "p-value notation: 0-0.001 = ***, 0.001-0.01 = **, 0.01-0.05 = *, 0.05-0.1 = .",
    notation = "none"
    ) %>%
  kable_styling(
    "striped",
    fixed_thead = TRUE,
    html_font = 'Consolas'
    ) %>%
pack_rows("Type of Residence", 12, 20) %>%
pack_rows("Cathedral Ceiling", 21, 22) %>%
pack_rows("Attic Finish", 23, 25) %>%
pack_rows("Garage 1 Attached", 26, 27) %>%
pack_rows("Garage 1 Area Included", 28, 29) %>%
pack_rows("Garage 1 Size", 30, 37) %>%
pack_rows("Garage 1 Ext. Wall Material", 38, 41) %>%
pack_rows("Attic Type", 42, 44) %>%
pack_rows("Basement Type", 45, 48) %>%
pack_rows("Exterior Wall Material", 49, 51) %>%
pack_rows("Central Heating", 52, 55) %>%
pack_rows("Basement Finish", 56, 58) %>%
pack_rows("Roof Material", 59, 64) %>%
pack_rows("Porch", 65, 66) %>%
pack_rows("Central Air Conditioning", 67, 68) %>%
pack_rows("Design Plan", 69, 70) %>%
  pack_rows("Recent Renovation", 71, 71)

logits <- logit_parcels %>%
  split(.$`Meta Township Name`) %>%
  sapply(function(x) {

    if (nrow(x) < 1000) {
      
      trim_vars <- grep("Renovation", vars, value = TRUE, invert = TRUE)
      
      fmla <- as.formula(
        paste0("Sale ~ ", paste0(gsub(" |-", "_", trim_vars), collapse = "+"))
        )
      
      glm(
      fmla,
      data = x %>%
        rename_with(~ gsub(" |-", "_", .x)), family = "binomial"
      ) %>%
  broom::tidy() %>%
  mutate(
    p.value = paste0(scales::pvalue(p.value), (stars.pval(p.value))),
    term = gsub("_", " ", term)
  ) %>%
  kable(
    col.names = c("Predictor", "β", "SE", "t", "p"),
    digits = c(0, 5, 5, 3, 0),
    align = c("l", "r", "c", "r", "l"),
    padding = 5L
  ) %>%
  add_footnote(
    "p-value notation: 0-0.001 = ***, 0.001-0.01 = **, 0.01-0.05 = *, 0.05-0.1 = .",
    notation = "none"
    ) %>%
  kable_styling(
    "striped",
    position = "left",
    fixed_thead = TRUE,
    html_font = 'Consolas'
    ) 
      
    } else {
      
      fmla <- as.formula(
        paste0("Sale ~ ", paste0(gsub(" |-", "_", vars), collapse = "+"))
        )
      
      glm(
      fmla,
      data = x %>%
        rename_with(~ gsub(" |-", "_", .x)), family = "binomial"
      ) %>%
  broom::tidy() %>%
  mutate(
    p.value = paste0(scales::pvalue(p.value), (stars.pval(p.value))),
    term = gsub("_", " ", term)
  ) %>%
  kable(
    col.names = c("Predictor", "β", "SE", "t", "p"),
    digits = c(0, 5, 5, 3, 0),
    align = c("l", "r", "c", "r", "l"),
    padding = 5L
  ) %>%
  add_footnote(
    "p-value notation: 0-0.001 = ***, 0.001-0.01 = **, 0.01-0.05 = *, 0.05-0.1 = .",
    notation = "none"
    ) %>%
  kable_styling(
    "striped",
    position = "left",
    fixed_thead = TRUE,
    html_font = 'Consolas'
    )
      
    }
    
  }, simplify = FALSE)
  
```

::: {.panel-tabset}

## County
```{r}
county_logit
```
## Barrington
```{r}
logits[["Barrington"]]
```
## Berwyn
```{r}
logits[["Berwyn"]]
```
## Bloom
```{r}
logits[["Bloom"]]
```
## Bremen
```{r}
logits[["Bremen"]]
```
## Calumet
```{r}
logits[["Calumet"]]
```
## Cicero
```{r}
logits[["Cicero"]]
```
## Elk Grove
```{r}
logits[["Elk Grove"]]
```
## Evanston
```{r}
logits[["Evanston"]]
```
## Hanover
```{r}
logits[["Hanover"]]
```
## Hyde Park
```{r}
logits[["Hyde Park"]]
```
## Jefferson
```{r}
logits[["Jefferson"]]
```
## Lake
```{r}
logits[["Lake"]]
```
## Lake View
```{r}
logits[["Lake View"]]
```
## Lemont
```{r}
logits[["Lemont"]]
```
## Leyden
```{r}
logits[["Leyden"]]
```
## Lyons
```{r}
logits[["Lyons"]]
```
## Maine
```{r}
logits[["Maine"]]
```
## New Trier
```{r}
logits[["New Trier"]]
```
## Niles
```{r}
logits[["Niles"]]
```
## North Chicago
```{r}
logits[["North Chicago"]]
```
## Northfield
```{r}
logits[["Northfield"]]
```
## Norwood Park
```{r}
logits[["Norwood Park"]]
```
## Oak Park
```{r}
logits[["Oak Park"]]
```
## Orland
```{r}
logits[["Orland"]]
```
## Palatine
```{r}
logits[["Palatine"]]
```
## Palos
```{r}
logits[["Palos"]]
```
## Proviso
```{r}
logits[["Proviso"]]
```
## Rich
```{r}
logits[["Rich"]]
```
## River Forest
```{r}
logits[["River Forest"]]
```
## Riverside
```{r}
logits[["Riverside"]]
```
## Rogers Park
```{r}
logits[["Rogers Park"]]
```
## Schaumburg
```{r}
logits[["Schaumburg"]]
```
## South Chicago
```{r}
logits[["South Chicago"]]
```
## Stickney
```{r}
logits[["Stickney"]]
```
## Thornton
```{r}
logits[["Thornton"]]
```
## West Chicago
```{r}
logits[["West Chicago"]]
```
## Wheeling
```{r}
logits[["Wheeling"]]
```
## Worth
```{r}
logits[["Worth"]]
```

:::

Across the township regressions we see two characteristics that tend to be significant and have large magnitudes: type of residence (stories), and the recent renovation indicator. For the South Tri, the recent renovation indicator is most significant in Bloom and Palos. What this implies for predicted AVs and desk review is hard to say - the recent reno indicator is clearly important, but difficult to update on the fly. Keep an eye out for older homes with unexpectedly high AVs next to homes that have just been renovated.

## Standardized mean differences

The standardized (mean) difference is a measure of distance between two group means in terms of one or more variables.

$$SMD = {\text{Difference in mean outcome between groups} \over \text{Standard deviation of outcome among participants}}$$

We're looking for SMDs greater than .1 or .05 for sensitive covariates.

```{r SMDs}

smd_table <- CreateTableOne(
  vars = vars %>% grep("Median", ., invert = TRUE, value = TRUE),
  strata = "Sale",
  data = sf_parcels,
  test = FALSE,
  includeNA = TRUE
)

print(smd_table, smd = TRUE)

```

## Empirical Cumulative Distributions

Looking for overlap/similar distributions. Numeric variables scaled and centered.

Numeric columns:

```{r ECDFs1}

grobs <- lapply(names(sf_parcels %>% select(where(is.numeric))), function(x) {

  data <- sf_parcels %>% select(Sale, any_of(x))
  
    ggplot(data, aes(
      scale(sf_parcels %>% pull(x)),
      color = Sale)
    ) +
      stat_ecdf(geom = "step") +
      labs(x = str_wrap(x, width = 25)) +
      theme_minimal() +
      theme(
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none"
      )

  })

for (i in seq(1, length(grobs), by = 6)) {
  
  if (i == 13) {
    
      grid.arrange(

  grobs = list(
    grobs[[i]], grobs[[i + 1]], grobs[[i + 2]], grobs[[i + 3]],
    rectGrob(gp = gpar(col = "white")),
    rectGrob(gp = gpar(col = "white"))

    ),
  ncol = 3,
  nrow = 2

)
    
} else {
    
      grid.arrange(

  grobs = list(
    grobs[[i]], grobs[[i + 1]], grobs[[i + 2]],
    grobs[[i + 3]], grobs[[i + 4]], grobs[[i + 5]]
    ),
  ncol = 3,
  nrow = 2

)
    
  }
  
}

```

Factor columns:

```{r ECDFs2}

grobs <- lapply(names(sf_parcels %>% select(where(is.factor) & any_of(vars))), function(x) {

    ggplot(sf_parcels, aes(
      as.numeric(sf_parcels %>% pull(x)),
      color = Sale)
    ) +
      stat_ecdf(geom = "step") +
      labs(x = str_wrap(x, width = 25)) +
      theme_minimal() +
      theme(
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none"
      )

  })

for (i in seq(1, length(grobs), by = 6)) {

    if (i == 13) {
    
      grid.arrange(

  grobs = list(
    grobs[[i]], grobs[[i + 1]], grobs[[i + 2]],
    grobs[[i + 3]], grobs[[i + 4]],
    rectGrob(gp = gpar(col = "white"))

    ),
  ncol = 3,
  nrow = 2

)
    
} else {  
      grid.arrange(

  grobs = list(
    grobs[[i]], grobs[[i + 1]], grobs[[i + 2]],
    grobs[[i + 3]], grobs[[i + 4]], grobs[[i + 5]]
    ),
  ncol = 3,
  nrow = 2

)
  
}

}

```

## Sales Trends

::: {.panel-tabset}

## Median Sale Price

``` {r Price}

ggplot(
  sales %>%
    group_by(meta_year, sv_is_outlier) %>%
    summarise(median_price = median(meta_sale_price)) %>%
    ungroup(),
  aes(x = meta_year, y = median_price, group = sv_is_outlier)
  ) +
  geom_line(aes(color = sv_is_outlier)) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(y = "Median Sale Price", x = "Year", title = "By Year and Outlier Status") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
        )

temp <- ggplot(
  sales %>%
    filter(sv_is_outlier == FALSE & meta_year >= '2019') %>%
    group_by(meta_year, meta_township_name, meta_triad_name) %>%
    summarise(median_price = median(meta_sale_price)) %>%
    group_by(meta_township_name, meta_triad_name) %>%
    mutate(median_price_change = (median_price / first(median_price, order_by = meta_year)) - 1),
  aes(x = meta_year, y = median_price_change, group = meta_township_name)
  ) +
  geom_line(aes(color = meta_township_name)) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(y = "Median Sale Price", x = "Year",
       title = "Price / 2019 Price, Outliers Excluded", subtitle = "By Tri, Township, and Year") +
  theme_minimal() +
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
    ) +
  facet_wrap(vars(meta_triad_name), scales = "free")

ggplotly(temp)

```

``` {r}

temp <- sales %>%
  filter(sv_is_outlier == FALSE) %>%
  mutate(time_sale_week_of_year = lubridate::week(meta_sale_date)) %>%
    group_by(time_sale_week_of_year, time_sale_year, meta_triad_name) %>%
  summarise(median_sale_price = median(meta_sale_price, na.rm = TRUE)) %>%
  split(.$meta_triad_name)
  
temp_graph <- function(x, title, limits = NULL) {
    
    x %>%
      ggplot() +
  geom_line(aes(x = time_sale_week_of_year, y = median_sale_price)) +
  labs(y = "Median Sale Price", x = "Week", title = paste(title, "Tri, By Year and Week"), subtitle = "Outliers excluded") +
  theme_minimal() +
  scale_y_continuous(labels = scales::dollar, limits = limits) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank()
        ) +
  facet_wrap(vars(time_sale_year))
    
  }
  
temp[["City"]] %>% temp_graph(title = "City")

temp[["North"]] %>% temp_graph(title = "North")

temp[["South"]] %>% temp_graph(title = "South", limits = c(100000, 300000))

```

## Number of Sales

``` {r Volume1}

ggplot(
  sales,
  aes(x = meta_year, group = sv_is_outlier)
  ) + 
  geom_bar(aes(fill = sv_is_outlier)) +
    geom_text(aes(
      label = format(after_stat(count), big.mark=",")
      ), stat = "count", vjust = -1) +
  labs(y = "Number of Sales", x = "Year", title = "By Year and Outlier Status") +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma, limits = c(0, 60000)) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
        )

ggplot(
  sales,
  aes(x = time_sale_month_of_year, group = sv_is_outlier)
  ) + 
  geom_bar(aes(fill = sv_is_outlier)) +
  labs(y = "Number of Sales", x = "Year", title = "By Month and Outlier Status") +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma, limits = c(0, 6000)) +
  scale_x_continuous(breaks = 1:12, labels = month.name) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust=1)
        ) +
  facet_wrap(vars(meta_year))
```


:::

## Outlier Breakdown

::: {.panel-tabset}

## Outlier Types

``` {r Outlier Types}

ggplot(
  sales %>%
    filter(sv_outlier_type != "Not outlier") %>%
    group_by(sv_outlier_type) %>%
    summarise(count = n()) %>%
    ungroup(),
  aes(x = reorder(sv_outlier_type, -count), y = count)
  ) +
  labs(y = "Number of Sales", x = "Outlier Types", title = "Count", subtitle = "by Type") +
  geom_bar(stat = "identity") +
    geom_text(aes(
      label = count
      ), vjust = -0.5) +
    ylim(0, 3300) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 60, hjust=1),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
    )

sales %>%
  group_by(meta_year, sv_outlier_type) %>%
  summarise(n = n()) %>%
  rename(Year = meta_year) %>%
  pivot_wider(id_cols = Year, names_from = sv_outlier_type, values_from = n) %>%
  kable() %>%
  kable_styling("striped")

```

## Class

``` {r Classes}

ggplot(
  sales %>%
    group_by(meta_class) %>%
  summarise(percent = mean(case_when(sv_is_outlier == TRUE ~ 1, TRUE ~ 0))) %>%
    ungroup(),
  aes(x = meta_class, y = percent)
  ) +
  labs(title = "Share of Outliers", subtitle = "by Class") +
  geom_bar(stat = "identity") +
    geom_text(aes(
      label = scales::percent(percent, accuracy = .1)
      ), vjust = -0.5) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format()) +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    )

```

## Township

``` {r Townships}

ggplot(
  sales %>%
    group_by(meta_township_name) %>%
  summarise(percent = mean(case_when(sv_is_outlier == TRUE ~ 1, TRUE ~ 0))) %>%
    ungroup(),
  aes(x = reorder(meta_township_name, -percent), y = percent)
  ) +
  labs(title = "Share of Outliers", subtitle = "by Township") +
  geom_bar(stat = "identity") +
    geom_text(aes(
      label = scales::percent(percent, accuracy = .1)
      ), hjust = -0.1, angle = 90) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 0.05)) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    )

```

## Price

``` {r}

sales %>%
  group_by(sv_is_outlier) %>%
  summarise(
    Min = min(meta_sale_price),
    Median = median(meta_sale_price),
    Mean = mean(meta_sale_price),
    Max = max(meta_sale_price),
    `Std dev` = sd(meta_sale_price)
    ) %>%
  rename("Sale category" = sv_is_outlier) %>%
  mutate(across(Min:`Std dev`, dollar)) %>%
  kable() %>%
  kable_styling(
    full_width = T
    )

ggplot(sales, aes(
      log(meta_sale_price),
      fill = sv_is_outlier)
    ) +
      geom_density(alpha = 0.5) +
      labs(title = "Cumulative Distribution") +
      theme_minimal() +
  labs(x = "log(price)") +
  scale_y_continuous(labels = scales::percent_format()) +
      theme(
        legend.position="bottom",
        legend.title = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank()
      )

```

``` {r fig.height=6}

ggplot(
  sales %>%
    group_by(sv_is_outlier, meta_township_name) %>%
    summarise(median_sale_price = median(meta_sale_price)) %>%
    ungroup() %>%
    pivot_wider(
      id_cols = meta_township_name,
      names_from = sv_is_outlier,
      values_from = median_sale_price
      ) %>%
    mutate(
      percent = `TRUE` / `FALSE`,
      above_below = ifelse(percent > 1, "Outlier > Not Outlier", "Not Outlier > Outlier"),
      above_below = factor(above_below, levels = c("Outlier > Not Outlier", "Not Outlier > Outlier"))
      ),
  aes(x = reorder(meta_township_name, percent), y = percent)
  ) +
  labs(
    title = "Median Outlier Price / Median Not outlier Price",
    subtitle = "by Township",
    fill = "Median Price"
    ) +
  geom_bar(stat = "identity", aes(fill = above_below)) +
  coord_flip() +
    geom_text(aes(
      label = round(percent, 2)
      ), hjust = -0.2) +
  scale_y_continuous(limits = c(0, 5.5)) +
  theme_minimal() +
  theme(,
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    )

```
## Price Per SQFT

``` {r ppsft}

sales %>%
  filter(!is.na(char_bldg_sf) & char_bldg_sf > 0) %>%
  mutate(ppsf = meta_sale_price / char_bldg_sf) %>%
  group_by(sv_is_outlier) %>%
  summarise(
    Min = min(ppsf),
    Median = median(ppsf),
    Mean = mean(ppsf),
    Max = max(ppsf),
    `Std dev` = sd(ppsf)
    ) %>%
  rename("Sale category" = sv_is_outlier) %>%
  mutate(across(Min:`Std dev`, dollar)) %>%
  kable() %>%
  kable_styling(
    full_width = T
    )

ggplot(
  sales %>%
  filter(!is.na(char_bldg_sf) & char_bldg_sf > 0) %>%
  mutate(ppsf = meta_sale_price / char_bldg_sf),
  aes(
      log(ppsf),
      fill = sv_is_outlier)
    ) +
      geom_density(alpha = 0.5) +
      labs(title = "Cumulative Distribution") +
      theme_minimal() +
  labs(x = "log(price per sqft)") +
  scale_y_continuous(labels = scales::percent_format()) +
      theme(
        legend.position="bottom",
        legend.title = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank()
      )

```

## Decile Breakout

``` {r Upper}

sales %>%
  arrange(meta_sale_price) %>%
  mutate(decile = ntile(meta_sale_price, 10)) %>%
  filter(decile == 10) %>%
  group_by(sv_is_outlier) %>%
  summarise(
    Min = min(meta_sale_price),
    Median = median(meta_sale_price),
    Mean = mean(meta_sale_price),
    Max = max(meta_sale_price),
    `Std dev` = sd(meta_sale_price)
    ) %>%
  rename("Sale category" = sv_is_outlier) %>%
  mutate(across(Min:`Std dev`, dollar)) %>%
  kable(caption = "Top Decile") %>%
  kable_styling(
    full_width = T
    )

ggplot(
  sales %>%
  arrange(meta_sale_price) %>%
  mutate(decile = ntile(meta_sale_price, 10)) %>%
    group_by(decile) %>%
  summarise(percent = mean(case_when(sv_is_outlier == TRUE ~ 1, TRUE ~ 0))) %>%
    ungroup(),
  aes(x = decile, y = percent)
  ) +
  labs(title = "Share of Outliers", x = "Quantile") +
  geom_bar(stat = "identity") +
  geom_bar(stat = "identity") +
    geom_text(aes(
      label = scales::percent(percent, accuracy = .1)
      ), hjust = -0.1, angle = 90) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 0.09), breaks = seq(0, 0.1, 0.02)) +
  scale_x_continuous(breaks = seq(1, 10, 1)) +
  theme(
    axis.title.y = element_blank()
    )

ggplot(
  sales %>%
  arrange(meta_sale_price) %>%
  mutate(decile = ntile(meta_sale_price, 10)) %>%
  filter(decile == 10 & sv_outlier_type != "Not outlier") %>%
    group_by(sv_outlier_type) %>%
    summarise(count = n()) %>%
    ungroup(),
  aes(x = reorder(sv_outlier_type, -count), y = count)
  ) +
  labs(y = "Number of Sales", x = "Outlier Types", title = "Outlier Count", subtitle = "by Type, Top Decile") +
  geom_bar(stat = "identity") +
    geom_text(aes(
      label = count
      ), vjust = -0.5) +
    ylim(0, 900) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 60, hjust=1),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
    )

```

:::

## Sale Outliers

::: {.panel-tabset}

## Sales View

``` {r}
temp_sales <- sales %>%
  left_join(
    ccao::town_dict %>% select(township_code, township_name, triad_name)
  ) %>%
  filter(class != "200" & triad_name == "South")

temp_sales %>%
  filter(sale_filter_is_outlier) %>%
  mutate(`Outlier Type` = case_when(
    sale_price_log10 > sale_filter_upper_limit ~ "High",
    sale_price_log10 < sale_filter_lower_limit ~ "Low",
    TRUE ~ NA_character_
  )) %>%
  group_by(class, `Outlier Type`, township_name) %>%
  summarise(
    `Min Sale Price` = min(sale_price, na.rm = TRUE),
    `Median Sale Price` = median(sale_price, na.rm = TRUE),
    `Max Sale Price` = max(sale_price, na.rm = TRUE),
    n = n()
  ) %>%
  left_join(
    temp_sales %>%
      filter(!sale_filter_is_outlier) %>%
      group_by(class, township_name) %>%
      summarise(`Median Non-outlier Sale Price` = median(sale_price, na.rm = TRUE))
  ) %>%
  mutate(across(contains("Sale"), dollar)) %>%
  dplyr::rename(Class = class, "Township Name" = township_name) %>%
  datatable(
    rownames = FALSE,
    height = "500px",
    options = list(
      columnDefs = list(
        list(className = "dt-center", targets = c(1:7))
      )
    )
  )
```
