```{r _comp_setup_script}
source("../_setup.R")
```

# Comparables

Each card in the **assessment set** has 20 comparable sales from the training data.
These comparables sales are extracted from the model structure and given a score
(0 - 100%) based on how closely they share the same terminal leaf nodes as the
target property. See [this
vignette](https://ccao-data.github.io/lightsnip/articles/finding-comps.html)
for more background on the algorithm.

## Neighborhood-Level Mean Comparison Score

The maps below show the _mean comparable score_ for PINs in a given neighborhood.
A higher mean comparable score suggests that the neighborhood has more similar /
homogeneous properties.

```{r _comp_neighborhood_means_coding}
base_data <- assessment_card %>%
  mutate(triad = ccao::town_get_triad(as.character(township_code))) %>%
  select(meta_pin, meta_nbhd_code, meta_class, triad, meta_card_num) %>%
  left_join(comp_df, by = c("meta_pin" = "pin", "meta_card_num" = "card"))

generate_plot <- function(df) {
  df_sum <- df %>%
    mutate(mean_comp = rowMeans(
      select(., starts_with("comp_score_")),
      na.rm = TRUE
    )) %>%
    group_by(meta_nbhd_code) %>%
    summarise(mean_comp = mean(mean_comp, na.rm = TRUE))

  ccao::nbhd_shp %>%
    filter(triad_code == run_triad_code) %>%
    left_join(df_sum, by = c("town_nbhd" = "meta_nbhd_code")) %>%
    st_as_sf() %>%
    ggplot() +
    geom_sf(aes(fill = mean_comp)) +
    scale_fill_distiller(
      palette = "YlGnBu",
      direction = 1,
      na.value = "grey90",
      labels = scales::percent_format()
    ) +
    theme_minimal() +
    labs(fill = "Mean\nComparable\nScore") +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.x = element_blank(),
      axis.ticks.y = element_blank()
    )
}

plots <- list()
plots$"Overall" <- base_data %>%
  filter(triad == run_triad_code) %>%
  generate_plot()
plots$"Single-Family" <- base_data %>%
  filter(!(meta_class %in% c("211", "212")), triad == run_triad_code) %>%
  generate_plot()
plots$"Multi-Family" <- base_data %>%
  filter(meta_class %in% c("211", "212"), triad == run_triad_code) %>%
  generate_plot()
```

::: {.panel-tabset}

```{r _comp_neighborhood_level_means, results='asis'}
iwalk(plots, ~ {
  cat("## ", .y, "\n\n")

  print(.x)

  cat("\n\n")
})
```

:::

## Map of Select Comparable Sets

This map shows a selection of PINs and their 20 comparable sales. Black dots
represent the target pin, while the colored dots represent the comparable sales.
Each color represents a group of comparables given a target PIN.

```{r _comp_map}
# Selection of random PINs that are useful for testing comps
report_pins <- c(
  "13253180150000", # Typical 211 2-flat in Logan Square, 2100 sqft
  "05174150240000", # Large 6 bed Georgian in Winnetka, 4600 sqft
  "14074220130000", # Uptown 205, 3 bed, 2.5 bath, 2000 sqft
  "20102120200000", # Bronzeville greystone, 211, 11 beds, 4100 sqft
  "20321100160000", # Auburn Gresham 203, 5 bed, 3 bath, 2400 sqft
  "20202080210000", # West Englewood 211, 6 bed, 2 bath post-fire, 2000 sqft
  "17203230280000", # Pilsen 211 2-flat, 4 bed, 2 bath, 1800 sqft
  "14321100320000", # Lincoln Park 206, 4 bed, 3 bath, 4000 sqft
  "16054270120000", # Austin 205, 6 bed, 2 bath, 2000 sqft
  "20142140140000" # Hyde Park greystone, 210, 5 beds, 4 baths, 3000 sqft
)

comp_df_filtered <- comp_df %>%
  filter(pin %in% report_pins) %>%
  select(-starts_with("comp_score_")) %>%
  tidyr::pivot_longer(
    starts_with("comp_pin_"),
    values_to = "comp_pin",
    names_to = "comp",
    names_pattern = "comp_pin_([0-9]*)"
  ) %>%
  left_join(
    comp_df %>%
      filter(pin %in% report_pins) %>%
      select(-starts_with("comp_pin_")) %>%
      tidyr::pivot_longer(
        starts_with("comp_score_"),
        values_to = "comp_score",
        names_to = "comp",
        names_pattern = "comp_score_([0-9]*)"
      ),
    by = c("pin", "card", "comp"),
    relationship = "one-to-one"
  ) %>%
  filter(card == min(card), .by = "pin") %>%
  mutate(type = "Comp.") %>%
  left_join(
    training_data %>%
      select(
        meta_pin, meta_card_num, meta_sale_price, meta_sale_date,
        loc_latitude, loc_longitude,
        meta_class, char_bldg_sf, char_yrblt, char_ext_wall
      ),
    by = c("comp_pin" = "meta_pin", "card" = "meta_card_num"),
    relationship = "many-to-many"
  ) %>%
  left_join(
    assessment_card %>%
      select(meta_pin, meta_card_num, pred_card_initial_fmv),
    by = c("comp_pin" = "meta_pin", "card" = "meta_card_num"),
    relationship = "many-to-many"
  ) %>%
  left_join(
    assessment_data %>%
      select(meta_pin, meta_card_num, meta_certified_tot),
    by = c("comp_pin" = "meta_pin", "card" = "meta_card_num"),
    relationship = "many-to-many"
  ) %>%
  mutate(
    pin_card = interaction(pin, card, sep = "_"),
    meta_certified_tot = meta_certified_tot * 10
  ) %>%
  ccao::vars_recode(cols = any_of(c("char_ext_wall"))) %>%
  filter(
    meta_sale_date == max(meta_sale_date),
    .by = c("pin", "card", "comp_pin")
  )

target_df <- comp_df %>%
  select(pin, card) %>%
  filter(pin %in% report_pins) %>%
  filter(card == min(card), .by = "pin") %>%
  mutate(type = "Target") %>%
  left_join(
    training_data %>%
      select(
        meta_pin, meta_card_num, meta_sale_price, meta_sale_date
      ),
    by = c("pin" = "meta_pin", "card" = "meta_card_num"),
    relationship = "many-to-many"
  ) %>%
  inner_join(
    assessment_card %>%
      filter(!is.na(loc_latitude) & !is.na(loc_longitude)) %>%
      select(
        meta_pin, meta_card_num, pred_card_initial_fmv,
        loc_latitude, loc_longitude,
        meta_class, char_bldg_sf, char_yrblt, char_ext_wall
      ),
    by = c("pin" = "meta_pin", "card" = "meta_card_num"),
    relationship = "many-to-many"
  ) %>%
  left_join(
    assessment_data %>%
      select(meta_pin, meta_card_num, meta_certified_tot),
    by = c("pin" = "meta_pin", "card" = "meta_card_num"),
    relationship = "many-to-many"
  ) %>%
  mutate(
    pin_card = interaction(pin, card, sep = "_"),
    meta_certified_tot = meta_certified_tot * 10
  ) %>%
  filter(
    meta_sale_date == max(meta_sale_date) | is.na(meta_sale_date),
    .by = c("pin", "card")
  )

comp_palette <- colorFactor(
  palette = "Set3",
  domain = comp_df_filtered$pin_card
)

comp_plot_data <- bind_rows(comp_df_filtered, target_df) %>%
  mutate(
    circle_color = ifelse(type == "Target", "black", comp_palette(pin_card)),
    comp_link = ifelse(
      !is.na(comp_pin),
      paste0(
        "<br>Comp PIN: ",
        "<a target='_blank' rel='noopener noreferrer'",
        "href='https://www.cookcountyassessor.com/pin/",
        comp_pin, "'>", comp_pin, "</a> "
      ),
      ""
    ),
    comp_score = ifelse(
      !is.na(comp_score),
      paste0("<br>Score: ", scales::percent(comp_score, accuracy = 0.01)),
      ""
    )
  )

leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    data = comp_plot_data,
    ~loc_longitude,
    ~loc_latitude,
    opacity = 1,
    fillOpacity = 1,
    radius = 2,
    color = ~circle_color,
    popup = ~ paste0(
      "<br>Target PIN: ",
      "<a target='_blank' rel='noopener noreferrer'",
      "href='https://www.cookcountyassessor.com/pin/",
      pin, "'>", pin, "</a>",
      comp_link,
      comp_score,
      "<hr>",
      "Certified FMV: ",
      scales::dollar(meta_certified_tot, accuracy = 1),
      "<br>Initial Card FMV: ",
      scales::dollar(pred_card_initial_fmv, accuracy = 1),
      "<hr>",
      "Sale Date: ", meta_sale_date,
      "<br>Sale Price: ", scales::dollar(meta_sale_price, accuracy = 1),
      "<hr>",
      "Class: ", meta_class,
      "<br>Bldg Sqft: ", scales::comma(char_bldg_sf),
      "<br>Year Built: ", char_yrblt,
      "<br>Ext. Wall: ", char_ext_wall
    )
  ) %>%
  setView(
    lng = mean(comp_plot_data$loc_longitude),
    lat = mean(comp_plot_data$loc_latitude),
    zoom = 10
  )
```

