```{r _input_setup_script}
source("../_setup.R")
```

# Data and Processing

This section primarily looks at data from the **training set**. It does three things:

- Checks for characteristic and data issues in the input data
- Checks that the input data is being processed, flagged, and filled as expected
- Checks that post-modeling and data post-processing steps are applied correctly

## Input Data Issues 

::: panel-tabset

### Large Sqft. 211s and 212s

Table of 211s and 212s which are in the 100th percentile by class and township.
These properties should likely be desk reviewed and/or valued manually.

```{r _input_large_211_212}
training_data %>%
  filter(
    meta_class %in% c("211", "212"),
    meta_triad_code == run_triad_code,
    !ind_pin_is_multicard
  ) %>%
  group_by(meta_class, meta_township_name) %>%
  mutate(
    sf_ntile = ntile(char_bldg_sf, 100),
    mean_sf = mean(char_bldg_sf, na.rm = TRUE)
  ) %>%
  filter(sf_ntile >= 100) %>%
  mutate(across(ends_with("_sf"), scales::comma)) %>%
  select(
    PIN = meta_pin,
    `Township Name` = meta_township_name,
    Class = meta_class,
    `Num. Apts.` = char_apts,
    `Mean Sqft. by Town and Class` = mean_sf,
    `Building Sqft` = char_bldg_sf
  ) %>%
  datatable(
    rownames = FALSE,
    options = list(
      columnDefs = list(
        list(
          className = "dt-right",
          targets = c(3:5)
        )
      )
    )
  )
```

### PINs with 0 or Missing Land Value

Table of PINs with `NULL` land square footage or a land square footage of 0.
Such values don't affect the predicted value of the property, but they may
receive incorrect building/land value splits.

```{r _input_pins_weird_land}
assessment_pin %>%
  filter(meta_triad_code == run_triad_code) %>%
  filter(
    pred_pin_final_fmv_land == 0 |
      is.infinite(pred_pin_final_fmv_land / pred_pin_final_fmv_round) |
      is.na(pred_pin_final_fmv_land / pred_pin_final_fmv_round)
  ) %>%
  mutate(township_name = ccao::town_convert(meta_township_code)) %>%
  select(
    `Year` = meta_year,
    `Township Name` = township_name,
    `PIN` = meta_pin,
    `Class` = meta_class,
    `Land %` = pred_pin_final_fmv_land / pred_pin_final_fmv_round,
    `Land FMV` = pred_pin_final_fmv_land,
    `Land Sqft.` = char_land_sf
  ) %>%
  datatable(rownames = FALSE)
```

:::

## Input Data Processing

::: panel-tabset

### Home Improvement Exemptions

Checks if Home Improvement Exemption characteristics are getting
accurately updated. In general, we expect numeric values to be additive
(`OG SQFT + HIE SQFT = NEW SQFT`) and categorical values to be
replacements (`OG AIR + HIE AIR = HIE AIR`).

```{r _input_hie_characteristics}
input_hie_pins <- training_data %>%
  filter(
    hie_num_active == 1,
    meta_triad_code == run_triad_code,
    meta_year == as.numeric(metadata$assessment_data_year)
  ) %>%
  pull(meta_pin) %>%
  unique()

input_hie_chars_df <- chars_data %>%
  filter(
    meta_year == as.numeric(metadata$assessment_data_year),
    meta_pin %in% input_hie_pins
  ) %>%
  inner_join(
    hie_data %>%
      filter(
        pin %in% input_hie_pins,
        year >= as.numeric(metadata$assessment_data_year) - 4,
        qu_sqft_bld > 0
      ) %>%
      select(-year),
    by = join_by(meta_pin == pin)
  ) %>%
  left_join(
    training_data %>% select(
      meta_pin, meta_year,
      train_sqft = char_bldg_sf,
      train_fbath = char_fbath, train_air = char_air
    ),
    by = join_by(meta_pin, meta_year)
  ) %>%
  select(
    PIN = meta_pin,
    `OG SQFT` = char_bldg_sf,
    `HIE SQFT` = qu_sqft_bld,
    `NEW SQFT` = train_sqft,
    `OG FBATH` = char_fbath,
    `HIE FBATH` = qu_full_bath,
    `NEW FBATH` = train_fbath,
    `OG AIR` = char_air,
    `HIE AIR` = qu_air,
    `NEW AIR` = train_air
  )

input_hie_chars_df %>%
  datatable(rownames = FALSE)
```

### Characteristic Filling

Checks if our time filling techniques are working as expected.
We expect point-in-time characteristics to be filled forward and backward in time
for most features i.e. 2019 flood factor records will cover both past and future
years in the records below.

```{r _input_characteristic_filling}
input_char_filling_df <- training_data %>%
  filter(
    !ind_pin_is_multicard,
    min(meta_sale_date) < lubridate::make_date(
      as.numeric(metadata$input_max_sale_year) - 3
    ),
    max(meta_sale_date) > lubridate::make_date(
      as.numeric(metadata$input_max_sale_year) - 3
    ),
    .by = meta_pin
  ) %>%
  filter(n() > 5, .by = meta_pin) %>%
  select(
    PIN = meta_pin,
    Year = meta_year,
    `Sale Date` = meta_sale_date,
    `FS Flood Factor (2019)` = loc_env_flood_fs_factor,
    `Walkability (2017)` = loc_access_cmap_walk_total_score,
    `GS Rating (2021)` = prox_avg_school_rating_in_half_mile
  ) %>%
  arrange(PIN, Year, `Sale Date`)

input_char_filling_df %>%
  datatable(rownames = FALSE)
```

### Map of Townhome Complex IDs

Checks if townhome complex ID construction in the `ingest` stage is working
as expected. The map shows 5 random complex IDs selected from each township,
along with complexes nearby the randomly selected ones. The map should show that
nearby PINs in the same area/complex receive different complex IDs if they have
significantly different characteristics.

```{r _input_complex_id_map, warning=FALSE}
# Self-join to find 5 random complex IDs from each township + their nearest
# neighboring complex IDs
input_complex_id_locs <- complex_id_data %>%
  left_join(
    assessment_card %>%
      select(meta_pin, meta_card_num, loc_longitude, loc_latitude),
    by = "meta_pin"
  ) %>%
  group_by(meta_complex_id) %>%
  summarise(
    mean_lat = mean(loc_latitude, na.rm = TRUE),
    mean_lon = mean(loc_longitude, na.rm = TRUE)
  ) %>%
  mutate(
    targ_lat_low = mean_lat - 0.001,
    targ_lat_high = mean_lat + 0.001,
    targ_lon_low = mean_lon - 0.001,
    targ_lon_high = mean_lon + 0.001
  )

input_complex_id_targets <- complex_id_data %>%
  mutate(
    township_name = ccao::town_convert(meta_township_code),
    triad = ccao::town_get_triad(meta_township_code, name = TRUE)
  ) %>%
  mutate(complex_id_cnt = n(), .by = meta_complex_id) %>%
  filter(triad == run_triad, complex_id_cnt >= 6) %>%
  distinct(township_name, meta_complex_id) %>%
  group_by(township_name) %>%
  slice_sample(n = 5) %>%
  ungroup() %>%
  select(meta_complex_id) %>%
  left_join(input_complex_id_locs, by = "meta_complex_id") %>%
  left_join(
    input_complex_id_locs %>% select(-starts_with("targ_")),
    by = join_by(
      between(y$mean_lat, x$targ_lat_low, x$targ_lat_high),
      between(y$mean_lon, x$targ_lon_low, x$targ_lon_high)
    )
  ) %>%
  distinct(meta_complex_id = meta_complex_id.y)

input_complex_id_map <- input_complex_id_targets %>%
  left_join(complex_id_data, by = "meta_complex_id") %>%
  left_join(
    chars_data %>%
      filter(meta_year == metadata$assessment_data_year) %>%
      select(
        meta_pin,
        loc_longitude,
        loc_latitude,
        char_rooms,
        char_bldg_sf,
        char_yrblt
      ),
    by = "meta_pin"
  ) %>%
  filter(!is.na(loc_latitude) & !is.na(loc_longitude))

input_complex_id_palette <- colorFactor(
  palette = "Set1",
  domain = unique(input_complex_id_map$meta_complex_id)
)

leaflet(input_complex_id_map) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    ~loc_longitude,
    ~loc_latitude,
    popup = ~ paste0(
      "Complex ID: ", as.character(meta_complex_id),
      "<hr>",
      "PIN: ",
      "<a target='_blank' rel='noopener noreferrer' ",
      "href='https://www.cookcountyassessor.com/pin/", meta_pin,
      "'>", meta_pin, "</a>",
      "<br>Rooms: ", char_rooms,
      "<br>Building SF: ", scales::comma(char_bldg_sf),
      "<br>Year Built: ", char_yrblt
    ),
    radius = 5,
    weight = 1,
    color = ~ input_complex_id_palette(meta_complex_id),
    opacity = 1,
    fillColor = ~ input_complex_id_palette(meta_complex_id),
    fillOpacity = 1
  )
```

:::

## Post-Modeling

::: panel-tabset

### Multi-Card and Multi-PIN Aggregation

This table uses test pins to check if multi-card parcels are correctly aggregated to the full PIN.
Look for consistency with past year's results. See the post-modeling section of the README documentation
for details on card-level aggregation.

```{r _input_multicard_sample}
input_multicard_test_pins <- tribble(
  ~"pin", ~"triad",
  "17321110470000", "south",
  "05174150240000", "south",
  "05213220250000", "south",
  "08121220400000", "south",
  "06334030310000", "south",
  "20142150160000", "city",
  "20143110110000", "city",
  "20143120360000", "city",
  "10313170010000", "city",
  "14331230380000", "city",
  "14293000710000", "city"
)

input_multicard_test_df <- assessment_card %>%
  inner_join(
    assessment_pin %>%
      filter(
        meta_pin %in% input_multicard_test_pins$pin,
        meta_triad_code == run_triad_code
      ) %>%
      select(meta_pin),
    by = "meta_pin"
  ) %>%
  arrange(meta_pin, meta_card_num) %>%
  select(
    PIN = meta_pin, Class = char_class, Card = meta_card_num,
    `Card %` = meta_card_pct_total_fmv, SQFT = char_bldg_sf,
    `Card Initial` = pred_card_initial_fmv,
    `Card Final` = pred_card_final_fmv
  ) %>%
  left_join(
    assessment_pin %>%
      select(
        PIN = meta_pin, Proration = meta_tieback_proration_rate,
        `PIN Far` = prior_far_tot,
        `PIN Near` = prior_near_tot,
        `PIN Initial` = pred_pin_initial_fmv,
        `PIN Final` = pred_pin_final_fmv_round
      ),
    by = c("PIN")
  ) %>%
  mutate(
    `Card %` = scales::percent(`Card %`, accuracy = 0.01),
    `SQFT` = scales::comma(`SQFT`),
    `Proration` = scales::percent(`Proration`, accuracy = 0.01),
    `Card Initial` = scales::dollar(`Card Initial`),
    `Card Final` = scales::dollar(`Card Final`),
    across(starts_with("PIN "), scales::dollar)
  )

input_multicard_test_df %>%
  datatable(
    rownames = FALSE,
    options = list(
      columnDefs = list(
        list(
          className = "dt-right",
          targets = c(
            "Card %", "SQFT", "Card Initial", "Card Final", "PIN Far",
            "PIN Near", "PIN Initial", "PIN Final", "Proration"
          )
        )
      )
    )
  )
```

### Land Valuation

Plot showing the post-modeling land valuation cap (usually 50%) in action. We
expect to see a spike at the cap. Properties above the cap are usually pro-rated
or multi-card PINs.

```{r, _input_land_valuation, out.width="100%"}
input_land_valuation <- assessment_pin %>%
  filter(meta_triad_code == run_triad_code) %>%
  mutate(land_pct = pred_pin_final_fmv_land / pred_pin_final_fmv_round) %>%
  select(
    meta_pin,
    land_pct,
    pred_pin_final_fmv_bldg,
    pred_pin_final_fmv_land,
    pred_pin_final_fmv_round
  )

input_land_valuation %>%
  ggplot() +
  geom_histogram(aes(x = land_pct)) +
  scale_x_continuous(
    name = "Land % of Total FMV",
    labels = scales::label_percent(),
    limits = c(0, 1)
  ) +
  labs(y = "Count") +
  theme_minimal()
```

### Townhome Complex Averaging

This takes a sample of 50 townhome PINs to see how far apart the initial
predicted values are from the final values (which are the average of the
complex). We want to ensure that the average for complexes is reasonable and
doesn't move the initial predictions too much.

```{r _input_townhomes_210_295}
assessment_pin %>%
  filter(!is.na(meta_complex_id) & meta_triad_code == run_triad_code) %>%
  group_by(meta_complex_id) %>%
  mutate(
    fmvpsf = pred_pin_final_fmv / char_total_bldg_sf
  ) %>%
  summarise(
    n = n(),
    "Min Initial FMV" = min(pred_pin_initial_fmv, na.rm = TRUE),
    "Max Initial FMV" = max(pred_pin_initial_fmv, na.rm = TRUE),
    "Initial % Range" =
      (`Max Initial FMV` - `Min Initial FMV`) / `Min Initial FMV`,
    "Mean Final FMV" = mean(pred_pin_final_fmv, na.rm = TRUE),
    "FMV per SQFT SD" = sd(fmvpsf, na.rm = TRUE)
  ) %>%
  slice_max(order_by = `Initial % Range`, n = 50) %>%
  mutate(
    across(contains("SQFT"), ~ round(.x, 1)),
    across(contains("FMV"), dollar),
    across(contains("%"), ~ percent(.x, accuracy = 0.1))
  ) %>%
  relocate(c(meta_complex_id, n)) %>%
  rename("Complex ID" = meta_complex_id, "Count" = n) %>%
  datatable(
    rownames = FALSE,
    options = list(
      columnDefs = list(
        list(
          className = "dt-right",
          targets = c(2:6)
        )
      )
    )
  )
```

:::


## Potential Output Issues

::: panel-tabset

### Card Values

```{r _outcomes_card_values}
assessment_card %>%
  summarize(
    "NA Initial FMV" = sum(is.na(pred_card_initial_fmv)),
    "Negative Initial FMV" = sum(pred_card_initial_fmv < 0, na.rm = TRUE),
    "NA Final FMV" = sum(is.na(pred_card_final_fmv)),
    "Negative Final FMV" = sum(pred_card_final_fmv < 0, na.rm = TRUE)
  ) %>%
  datatable(rownames = FALSE)
```

### PIN Values

```{r _outcomes_pin_values}
assessment_pin %>%
  summarize(
    "NA Initial FMV" = sum(is.na(pred_pin_initial_fmv)),
    "Negative Initial FMV" = sum(pred_pin_initial_fmv < 0, na.rm = TRUE),
    "NA Final FMV" = sum(is.na(pred_pin_final_fmv)),
    "Negative Final FMV" = sum(pred_pin_final_fmv < 0, na.rm = TRUE),
    "NA Final Land FMV" = sum(is.na(pred_pin_final_fmv_land)),
    "Negative Final Land FMV" = sum(pred_pin_final_fmv_land < 0, na.rm = TRUE),
    "NA Final Building FMV" = sum(is.na(pred_pin_final_fmv_bldg)),
    "Negative Final Building FMV" = sum(
      pred_pin_final_fmv_bldg < 0,
      na.rm = TRUE
    )
  ) %>%
  datatable(rownames = FALSE)
```

:::
