{{< include ../_setup.qmd >}}

# Outcomes

## Changes in Individual Assessed Value

This section looks at the year-over-year change in assessed values. It separates the data into individual observations, identifying the PINs in each neighborhood/township which grew at the fastest rate.

::: panel-tabset

### Map of Largest YoY Changes

This map caps the legend at a 300% to increase legibility. It plots the 10 largest year-over-year increases in assessed value by neighborhood.

```{r _outcomes_map_of_largest_change}
outcomes_map_big_changes <- assessment_pin %>%
  group_by(meta_nbhd_code) %>%
  left_join(ccao::nbhd_shp, by = c("meta_nbhd_code" = "town_nbhd")) %>%
  distinct(meta_pin, .keep_all = TRUE) %>%
  filter(!is.na(prior_near_yoy_change_pct)) %>%
  mutate(
    prior_near_yoy_change_pct = prior_near_yoy_change_pct * 100,
    neighborhood_median_change = median(prior_near_yoy_change_pct)
  ) %>%
  slice_max(prior_near_yoy_change_pct, n = 10) %>%
  mutate(capped_yoy_change = pmin(prior_near_yoy_change_pct, 300))

outcomes_map_nbhd_increases <- st_as_sf(outcomes_map_big_changes) %>%
  distinct(meta_nbhd_code, .keep_all = TRUE)

outcomes_map_color_palette <-
  colorNumeric(
    palette = "PuRd",
    domain = c(min(outcomes_map_big_changes$prior_near_yoy_change_pct), 300)
  )

leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = outcomes_map_nbhd_increases,
    fillColor = "white",
    weight = 1,
    color = "black",
    opacity = 1,
    fillOpacity = 0.5
  ) %>%
  addCircleMarkers(
    data = outcomes_map_big_changes,
    ~loc_longitude,
    ~loc_latitude,
    color = ~ outcomes_map_color_palette(capped_yoy_change),
    opacity = 0.7,
    fillOpacity = 0.7,
    radius = 1,
    popup = ~ paste0(
      "PIN: ", meta_pin,
      "<br>Class: ", meta_class,
      "<br><br>",
      "Neighborhood: ",
      meta_nbhd_code,
      "<br>Prior year (near) FMV: ",
      scales::dollar(prior_near_tot),
      "<br>Cur. year (rounded) FMV: ",
      scales::dollar(pred_pin_final_fmv_round),
      "<br>YoY Change: ",
      round(prior_near_yoy_change_pct, 2), "%"
    )
  ) %>%
  addLegend(
    "bottomright",
    pal = outcomes_map_color_palette,
    values = c(min(outcomes_map_big_changes$prior_near_yoy_change_pct), 300),
    title = "YoY Change",
    labFormat = labelFormat(suffix = "%")
  ) %>%
  setView(
    lng = mean(outcomes_map_big_changes$loc_longitude),
    lat = mean(outcomes_map_big_changes$loc_latitude),
    zoom = 10
  )
```

### Table of Largest YoY Changes

This table excludes any PINs which had a prior year assessed value of less than $25,000. It then identifies the 5 PINs in each neighborhood which had the largest year-over-year increase in assessed value.

```{r _outcomes_table_of_largest_change}
datatable(
  outcomes_map_big_changes %>%
    distinct(meta_pin, .keep_all = TRUE) %>%
    filter(prior_near_tot > 25000) %>%
    group_by(meta_township_code) %>%
    slice_max(order_by = prior_near_yoy_change_nom, n = 5) %>%
    ungroup() %>%
    left_join(
      select(chars_data, meta_pin, char_beds, char_rooms, char_fbath, meta_year),
      by = c("meta_pin", "meta_year")
    ) %>%
    distinct(meta_pin, .keep_all = TRUE) %>%
    mutate(
      Pin = meta_pin,
      Township = meta_township_code,
      Neighborhood = meta_nbhd_code,
      Class = meta_class,
      Address = loc_property_address,
      "Previous Value" = scales::dollar(prior_near_tot),
      "Predicted Value" = scales::dollar(pred_pin_final_fmv_round),
      "Nominal Change" = scales::dollar(prior_near_yoy_change_nom),
      "Percent Change" = scales::percent(
        round(prior_near_yoy_change_pct, 2) / 100
      ),
      "Recent Sale Date" = sale_recent_1_date,
      "Recent Sale Price" = scales::dollar(sale_recent_1_price),
      "Missing Critical Value" = flag_char_missing_critical_value,
      "Year Built" = char_yrblt,
      "Total Building SF" = scales::comma(char_total_bldg_sf),
      Bedrooms = char_beds,
      Rooms = char_rooms,
      "Full Bathrooms" = char_fbath,
      "Median Neighborhood Change" = scales::percent(
        round(neighborhood_median_change, 2) / 100
      )
    ) %>%
    select(
      Pin,
      Township,
      Neighborhood,
      Class,
      Address,
      `Previous Value`,
      `Predicted Value`,
      `Nominal Change`,
      `Percent Change`,
      `Recent Sale Date`,
      `Recent Sale Price`,
      `Missing Critical Value`,
      `Year Built`,
      `Total Building SF`,
      `Bedrooms`,
      `Rooms`,
      `Full Bathrooms`,
      `Median Neighborhood Change`
    ),
  options = list(pageLength = 10, scrollX = TRUE),
  filter = "top",
  rownames = FALSE
)
```

:::

## Median Rate of Change by Neighborhood

```{r _outcomes_map_of_neighborhood_change}
outcomes_map_nbhd_breaks <- c(-Inf, 0, 20, 40, 60, 80, 100, Inf)
outcomes_map_nbhd_labels <- c(
  "< 0%", "0% - 20%", "20% - 40%", "40% - 60%",
  "60% - 80%", "80% - 100%", "> 100%"
)

outcomes_map_nbhd_increases <- outcomes_map_nbhd_increases %>%
  mutate(
    change_category = cut(
      neighborhood_median_change,
      breaks = outcomes_map_nbhd_breaks,
      labels = outcomes_map_nbhd_labels,
      include.lowest = TRUE
    )
  )

outcomes_map_nbhd_palette <- colorFactor(
  palette = c("#e34a33", "#ffffff", RColorBrewer::brewer.pal(6, "BuPu")),
  domain = outcomes_map_nbhd_increases$change_category
)

leaflet(data = outcomes_map_nbhd_increases) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    fillColor = ~ outcomes_map_nbhd_palette(change_category),
    color = "black",
    weight = 1,
    opacity = 1,
    fillOpacity = 0.7,
    smoothFactor = 0.5,
    popup = ~ paste0(
      "Township: ", township_name,
      "<br>Neighborhood: ", meta_nbhd_code,
      "<br>Change: ", round(neighborhood_median_change, 2), "%"
    )
  ) %>%
  addLegend(
    position = "bottomright",
    pal = outcomes_map_nbhd_palette,
    values = ~change_category,
    title = "Median YoY Change",
    opacity = 1
  )
```
