{{< include ../_setup.qmd >}}

# Sales

## Sales Data

::: panel-tabset

## By Year

*NOTE: Outliers are removed*

```{r _sales_town_year}
township <- training_data %>%
  filter(!sv_is_outlier, meta_triad_name == run_triad) %>%
  mutate(meta_year = as.numeric(meta_year)) %>%
  summarise(sales = n(), .by = c(meta_township_name, meta_year)) %>%
  select(
    Sales = sales,
    Year = meta_year,
    Township = meta_township_name
  ) %>%
  ggplot() +
  geom_line(aes(x = Year, y = Sales, color = Township)) +
  scale_y_continuous(labels = scales::comma, breaks = seq(0, 6000, 1000)) +
  scale_x_continuous(breaks = seq(
    (as.numeric(params$year) - 9),
    as.numeric(params$year), 1
  )) +
  theme_minimal() +
  theme(axis.title.x = element_blank())

ggplotly(township)
```

## By Quarter

Sales by by quarter for the previous 4 years. The goal is to see how different
segments of the market are changing as a proportion in each tri-town. Sales are
binned into 4 different groups based on price.

*NOTE: Outliers are removed*

```{r _sales_town_bin1, fig.height=8, fig.width=7, out.width="100%"}
training_data %>%
  filter(
    !sv_is_outlier,
    meta_triad_name == run_triad,
    meta_year >= as.numeric(params$year) - 4
  ) %>%
  mutate(
    Bin = cut(
      meta_sale_price,
      breaks = c(1, 100000, 300000, 600000, 1000000, max(meta_sale_price)),
      labels = c(
        "$1 - $100,000",
        "$100,000 - $300,000",
        "$300,000 - $600,000",
        "$600,000 - $1,000,000",
        "$1,000,000+"
      )
    ),
    Quarter = lubridate::quarter(meta_sale_date) + 4 *
      (as.numeric(meta_year) - as.numeric(params$year))
  ) %>%
  summarise(Sales = n(), .by = c(meta_township_name, Quarter, Bin)) %>%
  select(Sales, Bin, Township = meta_township_name, Quarter) %>%
  ggplot(aes(
    x = Quarter,
    y = Sales,
    fill = Bin,
    group = Bin
  )) +
  geom_area() +
  scale_color_brewer(palette = "PuOr") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  ) +
  facet_wrap(vars(Township), scales = "free_y", ncol = 3)
```

## By Class and Year

*NOTE: Outliers are removed*

```{r _sales_class_year}
class <- training_data %>%
  filter(
    !sv_is_outlier,
    meta_triad_name == run_triad
  ) %>%
  mutate(meta_year = as.numeric(meta_year)) %>%
  summarise(sales = n(), .by = c(meta_class, meta_year)) %>%
  rename(Class = meta_class) %>%
  pivot_wider(
    id_cols = Class,
    names_from = meta_year,
    values_from = sales
  )

class %>%
  kable() %>%
  kable_styling() %>%
  row_spec(unique(which(is.na(class), arr.ind = TRUE)[, 1]),
    bold = TRUE,
    background = "lightgrey"
  )
```

## Removed Outliers

These are outliers that were removed via the sale validation process in the
ingest stage. The goal is to confirm that these were reasonable sales to remove.

```{r _sales_removed_outliers}
rbind(
  training_data %>%
    filter(
      sv_is_outlier,
      meta_triad_name == run_triad,
      !ind_pin_is_multicard
    ) %>%
    slice_max(order_by = meta_sale_price, n = 25),
  training_data %>%
    filter(
      sv_is_outlier,
      meta_triad_name == run_triad,
      !ind_pin_is_multicard
    ) %>%
    slice_min(order_by = meta_sale_price, n = 25)
) %>%
  select(
    meta_sale_price:meta_tieback_proration_rate,
    loc_property_address:loc_property_zip,
    char_yrblt:char_beds,
    starts_with("loc_school"),
    sv_is_outlier,
    sv_outlier_type
  ) %>%
  datatable(rownames = FALSE)
```

## Kept Outliers

These are the top and bottom 25 sales from the training data that are *still* in
the sample. The goal is to gauge whether or not outlier removal is being too
strict (or not strict enough).

```{r _sales_kept_outliers}
rbind(
  training_data %>%
    filter(
      !sv_is_outlier,
      meta_triad_name == run_triad,
      !ind_pin_is_multicard
    ) %>%
    slice_max(order_by = meta_sale_price, n = 25),
  training_data %>%
    filter(
      !sv_is_outlier,
      meta_triad_name == run_triad,
      !ind_pin_is_multicard
    ) %>%
    slice_min(order_by = meta_sale_price, n = 25)
) %>%
  select(
    meta_sale_price:meta_tieback_proration_rate,
    loc_property_address:loc_property_zip,
    char_yrblt:char_beds,
    starts_with("loc_school"),
    sv_is_outlier,
    sv_outlier_type
  ) %>%
  datatable(rownames = FALSE)
```

:::

## Sales Trends

::: panel-tabset

## Median Sale Price by Year and Outlier Status

```{r _sales_median_price}
ggplot(
  training_data %>%
    summarise(
      median_price = median(meta_sale_price),
      .by = c(meta_year, sv_is_outlier)
    ),
  aes(x = meta_year, y = median_price, group = sv_is_outlier)
) +
  geom_line(aes(color = sv_is_outlier)) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_discrete(name = "Outlier Status", labels = c("Not Outlier", "Outlier")) +
  labs(
    y = "Median Sale Price", x = "Year"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )


price_change <- ggplot(
  training_data %>%
    filter(sv_is_outlier == FALSE &
      meta_year >= as.numeric(params$year) - 4) %>%
    group_by(meta_year, meta_township_name, meta_triad_name) %>%
    summarise(median_price = median(meta_sale_price)) %>%
    group_by(meta_township_name, meta_triad_name) %>%
    mutate(median_price_change = (
      median_price /
        first(median_price, order_by = meta_year)
    ) - 1),
  aes(x = meta_year, y = median_price_change, group = meta_township_name)
) +
  geom_line(aes(color = meta_township_name)) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    y = "Median Sale Price",
    x = "Year",
    title = paste(
      "Price / Price of", as.numeric(params$year) - 4, ", Outliers Excluded"
    ),
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  ) +
  facet_wrap(vars(meta_triad_name), scales = "free")

ggplotly(price_change)
```

## Sale Price by Year and Week
```{r _sales_price_week}
tri_sales <- training_data %>%
  filter(sv_is_outlier == FALSE) %>%
  mutate(time_sale_week_of_year = lubridate::week(meta_sale_date)) %>%
  group_by(time_sale_week_of_year, time_sale_year, meta_triad_name) %>%
  summarise(median_sale_price = median(meta_sale_price, na.rm = TRUE)) %>%
  split(.$meta_triad_name)


plots <- map2(tri_sales, names(tri_sales), function(x, title) {
  limits <- if (title == "South") c(100000, 300000) else NULL

  x %>%
    ggplot() +
    geom_line(aes(x = time_sale_week_of_year, y = median_sale_price)) +
    labs(
      y = "Median Sale Price",
      x = "Week",
      title = paste(title, "Tri"),
      subtitle = "Outliers excluded"
    ) +
    theme_minimal() +
    scale_y_continuous(labels = scales::dollar, limits = limits) +
    theme(
      legend.position = "bottom",
      legend.title = element_blank(),
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x = element_blank()
    ) +
    facet_wrap(vars(time_sale_year))
})

extract_and_print_plots <- function(plots) {
  num_plots <- length(plots)

  for (i in 1:num_plots) {
    print(plots[[i]])
  }
}

extract_and_print_plots(plots)
```

## Sales by Year, Month, and Outlier Status

```{r _sales_volume_by_year_month_outlier}
ggplot(
  training_data,
  aes(x = meta_year, group = sv_is_outlier)
) +
  geom_bar(aes(fill = sv_is_outlier)) +
  geom_text(aes(
    label = format(after_stat(count), big.mark = ",")
  ), stat = "count", vjust = -1) +
  labs(
    y = "Number of Sales", x = "Year"
  ) +
  scale_fill_discrete(name = "Outlier Status", labels = c("Not Outlier", "Outlier")) +
  theme_minimal() +
  scale_y_continuous(
    labels = comma,
    limits = c(0, 60000)
  ) +
  theme(
    legend.position = "bottom",
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )

ggplot(
  training_data,
  aes(x = time_sale_month_of_year, group = sv_is_outlier)
) +
  geom_bar(aes(fill = sv_is_outlier)) +
  labs(
    y = "Number of Sales", x = "Year"
  ) +
  scale_fill_discrete(name = "Outlier Status", labels = c("Not Outlier", "Outlier")) +
  theme_minimal() +
  scale_y_continuous(labels = comma, limits = c(0, 6000)) +
  scale_x_continuous(breaks = 1:12, labels = month.name) +
  theme(
    legend.position = "bottom",
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  facet_wrap(vars(meta_year))
```

:::
