```{r _stats_setup_script}
source("../_setup.R")
```

# Statistical Tests

This section contains various statistical tests and plots, mostly to ensure that the universe of sold properties is representative of the universe of all properties.

```{r _stats_gather_data}
stats_regression_vars <-
  grep("char_|3435|acs5_median|township_name",
    names(chars_data),
    value = TRUE
  ) %>%
  grep("percentile|_apts|_ncu|_use|_qlty",
    .,
    value = TRUE,
    invert = TRUE
  )

# All residential parcel characteristics from time frame, with a sale indicator
stats_sf_parcels <- chars_data %>%
  filter(!meta_class %in% c("211", "212")) %>%
  mutate(meta_card_num = as.character(meta_card_num)) %>%
  # Join on sales indicator for most recent sale in the last two years -
  # if there is one
  left_join(
    training_data %>%
      filter(
        year >= as.numeric(metadata$input_max_sale_year) - 1,
        !sv_is_outlier, !ind_pin_is_multicard,
        meta_sale_date == max(meta_sale_date, na.rm = TRUE),
        .by = c(meta_pin, meta_card_num)
      ) %>%
      # We apply distinct here to remove duplicate sales that are a result of
      # multicard and multiland sales, otherwise we'll get many-to-many joins.
      distinct() %>%
      select(meta_pin, meta_card_num, meta_sale_date, meta_year)
  ) %>%
  mutate(
    sale = factor(
      as.numeric(!is.na(meta_sale_date)),
      levels = c(0, 1),
      labels = c("No Sale", "Sale")
    ),
    char_recent_renovation = case_when(
      char_recent_renovation == 0 ~ "No",
      TRUE ~ "Yes"
    ),
    across(
      (starts_with("char_") & where(is.character)) |
        c(char_recent_renovation, char_type_resd),
      as.factor
    )
  ) %>%
  select(sale, all_of(stats_regression_vars)) %>%
  vars_recode(
    code_type = "long",
    dictionary = ccao::vars_dict_legacy
  ) %>%
  vars_rename(names_from = "athena", names_to = "pretty") %>%
  rename_with(~ gsub("loc_", "", .x)) %>%
  rename_with(~ gsub("_", " ", .x)) %>%
  rename_with(toTitleCase) %>%
  rename_with(~ gsub("Acs5", "ACS5", .x))

stats_logit_parcels <- stats_sf_parcels %>%
  mutate(across(where(is.factor), ~ addNA(.x))) %>%
  drop_na()

stats_logit_vars <- names(stats_sf_parcels) %>%
  grep("Sale|Township", ., value = TRUE, invert = TRUE)
```

## Logistic Regression

All covariates regressed on sale status - significance indicates a non-random effect on the likelihood of a sale, ceterus paribus. Sample is universe of single-family parcels from the two years prior to assessment, excluding some parcels with incomplete characteristics.

```{r _stats_logit}
stats_logit_remove <- stats_sf_parcels %>%
  rename_with(~ gsub(" |-", "_", .x)) %>%
  select(where(is.factor)) %>%
  names() %>%
  grep(
    "Sale|Township|Apartments|Single|Commercial",
    .,
    invert = TRUE,
    value = TRUE
  )

stats_logit_fmla <- as.formula(
  paste0("Sale ~ ", paste(gsub(" |-", "_", stats_logit_vars), collapse = "+"))
)

stats_logit_fit <- glm(stats_logit_fmla,
  data = stats_logit_parcels %>%
    rename_with(~ gsub(" |-", "_", .x)),
  family = "binomial"
)

stats_logit_sample_size <- stats_logit_fit$df.null

stats_county_logit <- stats_logit_fit %>%
  broom::tidy() %>%
  mutate(
    p.value = paste0(scales::pvalue(p.value), (stars.pval(p.value))),
    term = gsub(
      "_",
      " ",
      gsub(paste(stats_logit_remove, collapse = "|"), "", term)
    )
  ) %>%
  kable(
    caption = paste0(
      "Sample size: ",
      format(stats_logit_sample_size, big.mark = ","),
      " single-family parcels"
    ),
    col.names = c("Predictor", "β", "SE", "t", "p"),
    digits = c(0, 5, 5, 3, 0),
    align = c("l", "r", "c", "r", "l"),
    padding = 5L
  ) %>%
  add_footnote(
    paste0(
      "p-value notation: 0-0.001 = ***, 0.001-0.01 = **, ",
      "0.01-0.05 = *, 0.05-0.1 = ."
    ),
    notation = "none"
  ) %>%
  kable_styling("striped",
    fixed_thead = TRUE,
    html_font = "Consolas"
  ) %>%
  pack_rows("Type of Residence", 12, 20) %>%
  pack_rows("Cathedral Ceiling", 21, 22) %>%
  pack_rows("Attic Finish", 23, 25) %>%
  pack_rows("Garage 1 Attached", 26, 27) %>%
  pack_rows("Garage 1 Area Included", 28, 29) %>%
  pack_rows("Garage 1 Size", 30, 37) %>%
  pack_rows("Garage 1 Ext. Wall Material", 38, 41) %>%
  pack_rows("Attic Type", 42, 44) %>%
  pack_rows("Basement Type", 45, 48) %>%
  pack_rows("Exterior Wall Material", 49, 51) %>%
  pack_rows("Central Heating", 52, 55) %>%
  pack_rows("Basement Finish", 56, 58) %>%
  pack_rows("Roof Material", 59, 64) %>%
  pack_rows("Porch", 65, 66) %>%
  pack_rows("Central Air Conditioning", 67, 68) %>%
  pack_rows("Design Plan", 69, 70) %>%
  pack_rows("Recent Renovation", 71, 71)
```

```{r _stats_county_regression}
stats_county_logit
```

## Standardized Mean Differences

The standardized (mean) difference is a measure of distance between two group
means in terms of one or more variables.

$$SMD = {\text{Difference in mean outcome between groups} \over \text{Standard
deviation of outcome among participants}}$$

We're looking for SMDs greater than .1 or .05 for sensitive covariates.

```{r _stats_smds}
stats_smd_table <- CreateTableOne(
  vars = stats_logit_vars %>%
    grep("Median", ., invert = TRUE, value = TRUE),
  strata = "Sale",
  data = stats_sf_parcels,
  test = FALSE,
  includeNA = TRUE
)

tableone::kableone(stats_smd_table, smd = TRUE)
```

## Empirical Cumulative Distributions

For the empirical cumulative distributions, we are looking for overlap/similar
distributions. Numeric variables scaled and centered.

::: panel-tabset

### Numeric Columns

```{r _stats_ecdfs_numeric}
stats_ecdfs_plots_num <- stats_sf_parcels %>%
  select(where(is.numeric)) %>%
  length()

stats_ecdfs_grobs_num <- lapply(seq_len(stats_ecdfs_plots_num), function(i) {
  x <- names(stats_sf_parcels %>% select(where(is.numeric)))[i]
  data <- stats_sf_parcels %>% select(Sale, any_of(x))
  ggplot(data, aes(scale(stats_sf_parcels %>% pull(x)), color = Sale)) +
    stat_ecdf(geom = "step") +
    labs(x = str_wrap(x, width = 25)) +
    theme_minimal() +
    theme(
      axis.title.y = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      legend.position = if (i == stats_ecdfs_plots_num) "bottom" else "none"
    )
})

for (i in seq(1, length(stats_ecdfs_grobs_num), by = 6)) {
  if (i == 13) {
    grid.arrange(
      grobs = list(
        stats_ecdfs_grobs_num[[i]],
        stats_ecdfs_grobs_num[[i + 1]],
        stats_ecdfs_grobs_num[[i + 2]],
        stats_ecdfs_grobs_num[[i + 3]],
        rectGrob(gp = gpar(col = "white")),
        rectGrob(gp = gpar(col = "white"))
      ),
      ncol = 3,
      nrow = 2
    )
  } else {
    grid.arrange(
      grobs = list(
        stats_ecdfs_grobs_num[[i]],
        stats_ecdfs_grobs_num[[i + 1]],
        stats_ecdfs_grobs_num[[i + 2]],
        stats_ecdfs_grobs_num[[i + 3]],
        stats_ecdfs_grobs_num[[i + 4]],
        stats_ecdfs_grobs_num[[i + 5]]
      ),
      ncol = 3,
      nrow = 2
    )
  }
}
```

### Factor Columns

```{r _stats_ecdfs_factor}
stats_ecdfs_plots_fct <- stats_sf_parcels %>%
  select(where(is.factor), -contains("class")) %>%
  length()

stats_ecdfs_grobs_fct <- lapply(seq_len(stats_ecdfs_plots_fct), function(i) {
  x <- names(
    stats_sf_parcels %>%
      select(where(is.factor) & any_of(stats_logit_vars))
  )[i]

  ggplot(
    data = stats_sf_parcels,
    mapping = aes(as.numeric(stats_sf_parcels %>% pull(x)), color = Sale)
  ) +
    stat_ecdf(geom = "step") +
    labs(x = str_wrap(x, width = 25)) +
    theme_minimal() +
    theme(
      axis.title.y = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      legend.position = if (i == stats_ecdfs_plots_fct) "bottom" else "none"
    )
})

for (i in seq(1, length(stats_ecdfs_grobs_fct), by = 6)) {
  if (i == 13) {
    grid.arrange(
      grobs = list(
        stats_ecdfs_grobs_fct[[i]],
        stats_ecdfs_grobs_fct[[i + 1]],
        stats_ecdfs_grobs_fct[[i + 2]],
        stats_ecdfs_grobs_fct[[i + 3]],
        stats_ecdfs_grobs_fct[[i + 4]],
        rectGrob(gp = gpar(col = "white"))
      ),
      ncol = 3,
      nrow = 2
    )
  } else {
    grid.arrange(
      grobs = list(
        stats_ecdfs_grobs_fct[[i]],
        stats_ecdfs_grobs_fct[[i + 1]],
        stats_ecdfs_grobs_fct[[i + 2]],
        stats_ecdfs_grobs_fct[[i + 3]],
        stats_ecdfs_grobs_fct[[i + 4]],
        stats_ecdfs_grobs_fct[[i + 5]]
      ),
      ncol = 3,
      nrow = 2
    )
  }
}
```

:::

## Change In and Out-of-Sample

If the training sample is representative of the housing stock at large we should
see similar changes in AVs between the previous year and predictions regardless
of whether a home has sold recently. This table compares the median change in AV
(dollar and percentage) for homes which have sold in the three years prior to
the assessment year those which did not sell.

```{r _stats_in_v_out_of_sample_delta}
stats_median_yoy_delta <- assessment_pin %>%
  mutate(
    Township = ccao::town_convert(meta_township_code),
    prior_far_yoy_change_pct = as.numeric(prior_far_yoy_change_pct),
    prior_near_yoy_change_nom = as.numeric(prior_near_yoy_change_nom),
    is_recent_sale = case_when(
      lubridate::year(sale_recent_1_date) >=
        as.numeric(metadata$input_max_sale_year) - 3 ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  filter(
    !is.na(prior_near_yoy_change_pct),
    !is.na(prior_near_yoy_change_nom)
  ) %>%
  summarize(
    yoy_change_percent = median(prior_near_yoy_change_pct, na.rm = TRUE),
    yoy_change_dollars = median(prior_near_yoy_change_nom, na.rm = TRUE),
    .by = c(Township, is_recent_sale)
  ) %>%
  pivot_wider(
    id_cols = Township,
    names_from = is_recent_sale,
    values_from = c(yoy_change_percent, yoy_change_dollars)
  ) %>%
  mutate(
    change_percent_ratio =
      round(yoy_change_percent_TRUE / yoy_change_percent_FALSE, 2),
    change_dollars_ratio =
      round(yoy_change_dollars_TRUE / yoy_change_dollars_FALSE, 2),
    yoy_change_percent_TRUE =
      scales::percent(yoy_change_percent_TRUE, accuracy = 2),
    yoy_change_percent_FALSE =
      scales::percent(yoy_change_percent_FALSE, accuracy = 2),
    yoy_change_dollars_TRUE = scales::dollar(yoy_change_dollars_TRUE),
    yoy_change_dollars_FALSE = scales::dollar(yoy_change_dollars_FALSE)
  ) %>%
  select(
    Township,
    change_percent_ratio, yoy_change_percent_TRUE, yoy_change_percent_FALSE,
    change_dollars_ratio, yoy_change_dollars_TRUE, yoy_change_dollars_FALSE,
  )

stats_median_yoy_delta %>%
  mutate(
    change_percent_ratio = cell_spec(
      change_percent_ratio,
      color = case_when(
        is.na(change_percent_ratio) | is.nan(change_percent_ratio) ~ "grey80",
        between(change_percent_ratio, 0.95, 1.05) ~ "green",
        TRUE ~ "red"
      )
    ),
    change_dollars_ratio = cell_spec(
      change_dollars_ratio,
      color = case_when(
        is.na(change_dollars_ratio) | is.nan(change_dollars_ratio) ~ "grey80",
        between(change_dollars_ratio, 0.95, 1.05) ~ "green",
        TRUE ~ "red"
      )
    )
  ) %>%
  kable(
    "html",
    escape = FALSE, col.names = rep("", 7), align = c("l", rep("c", 6))
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(
    "Township" = 1,
    "Ratio" = 1,
    "Sold" = 1,
    "Unsold" = 1,
    "Ratio" = 1,
    "Sold" = 1,
    "Unsold" = 1
  )) %>%
  add_header_above(c(" " = 1, "Percentage Change" = 3, "Dollar Change" = 3)) %>%
  add_header_above(
    c("Comparison of YOY Change in AV for Sold and Unsold Houses" = 7)
  )
```
