# Triad-Level Machine Learning Metrics

This section compares model performance metrics at the triad level between the
**training set** (the data the model learned from) and the **test set** (a
held-out sample the model never saw during training). Large gaps between train
and test metrics — particularly for RMSE, MDAPE, or MAE — may indicate
overfitting, where the model has memorized the training data rather than
learning generalizable patterns. The sale price and prediction distributions
are included to confirm that both splits are drawn from similar underlying
price ranges.

**Note**: Consistent with the rest of this report, while the model was trained
on county-wide data, we report these splits at the triad level since that is
where we are most concerned with overfitting.

```{r _train_test_setup_script}
source("../_setup.R")
```

```{r _train_test_pull_stats}
test_stats <- model_performance_test %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_stats <- model_performance_train %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_test_comparison_tbl <- dplyr::bind_rows(list(
  "train" = train_stats, "test" = test_stats
), .id = "split")
```


```{r _train_test_construct_table}
train_test_comparison_tbl %>%
  select(
    split, rmse, mdape, mae,
    sale_fmv_mean, sale_fmv_sd,
    estimate_fmv_mean, estimate_fmv_sd
  ) %>%
  mutate(across(c(
    rmse, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  ), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Train vs. Test Set Performance Metrics",
    col.names = c(
      "Split", "RMSE", "MdAPE", "MAE",
      "Mean Sale Price", "SD Sale Price",
      "Mean Estimate", "SD Estimate"
    )
  )
```

```{r _train_test_actual_price_distributions}
train_test_comparison_tbl %>%
  select(
    split,
    sale_fmv_min, sale_fmv_q25, sale_fmv_median, sale_fmv_q75, sale_fmv_max
  ) %>%
  mutate(across(starts_with("sale_fmv_"), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Actual Sale Price Distributions",
    col.names = c("Split", "Min", "Q1", "Median", "Q3", "Max")
  )
```

```{r _train_test_estimated_distributions}
train_test_comparison_tbl %>%
  select(
    split,
    estimate_fmv_min, estimate_fmv_q25, estimate_fmv_median,
    estimate_fmv_q75, estimate_fmv_max
  ) %>%
  mutate(across(starts_with("estimate_fmv_"), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Predicted Sale Price Distributions",
    col.names = c("Split", "Min", "Q1", "Median", "Q3", "Max")
  )
```

```{r _train_test_metrics}
# This provides the count of PINs in each township for the modeling year.
# We want to see if the stats differ for the test set when there are
# a low percentage of the housing stock being sold.
housing_stock <- assessment_pin %>%
  # Filter to modeling triad
  filter(meta_triad_code == run_triad_code) %>%
  group_by(meta_township_code) %>%
  summarise(total_units = n_distinct(meta_pin), .groups = "drop") %>%
  mutate(meta_township_code = as.character(meta_township_code))

# Function to produce one metric for individual townships in the test
# set and aggregate metrics for the test and training set.
calc_metrics <- function(df, pred_col, group_vars = NULL) {
  group_vars <- group_vars %||% c(
    "month_year", "township", "meta_township_code", "total_units"
  )

  df <- df %>%
    filter(meta_triad_code == run_triad_code) %>%
    mutate(
      meta_township_code = as.character(meta_township_code),
      township           = ccao::town_convert(meta_township_code),
      .pred              = .data[[pred_col]],
      ratio              = .pred / meta_sale_price,
      month_year         = floor_date(as.Date(meta_sale_date), "month")
    )

  if (any(c("township", "total_units") %in% group_vars)) {
    df <- left_join(df, housing_stock, by = "meta_township_code")
  }

  df %>%
    group_by(across(all_of(group_vars))) %>%
    summarise(
      median_ratio = median(ratio, na.rm = TRUE),
      rmse = sqrt(mean((.pred - meta_sale_price)^2,
        na.rm = TRUE
      )),
      mape = mean(
        abs((.pred - meta_sale_price) / meta_sale_price),
        na.rm = TRUE
      ) * 100,
      n = n(),
      .groups = "drop"
    ) %>%
    mutate(
      month_num = as.numeric(month_year),
      volume_pct = if ("total_units" %in% names(.)) {
        (n / total_units) * 100
      } else {
        NA_real_
      }
    )
}

# The chart becomes cluttered with training metrics for each
# township so we just produce an overarching train metric
# and then test metrics for each township if anyone wants
# a deeper dive.

metrics_test <- calc_metrics(
  test_card,
  "pred_card_initial_fmv"
)
total_test <- calc_metrics(test_card,
  "pred_card_initial_fmv",
  group_vars = "month_year"
)
total_train <- calc_metrics(
  train_card, "pred_card_initial_fmv",
  group_vars = "month_year"
)

ratio_line <- list(
  type = "line", xref = "paper", x0 = 0, x1 = 1,
  y0 = 1, y1 = 1, line = list(color = "red", width = 1.5, dash = "dash")
)

# Add a scatter trace + optional loess trend trace
add_metric_traces <- function(p, data, metric_col, x_col, name, color,
                              mode = "lines+markers", marker = NULL,
                              line = NULL, legendgroup = name,
                              visible = TRUE, span = 0.75) {
  d_fit <- filter(data, !is.na(.data[[metric_col]]))

  p <- p %>% add_trace(
    data = data, x = as.formula(paste0("~", x_col)),
    y = as.formula(paste0("~", metric_col)),
    name = name, type = "scatter", mode = mode,
    legendgroup = legendgroup,
    line = line %||% list(color = color, width = 2),
    marker = marker %||% list(color = color, size = 5),
    visible = visible
  )

  if (nrow(d_fit) >= 3) {
    # Use numeric month for loess, then assign trend back
    d_fit$trend <- predict(loess(
      as.formula(paste0(metric_col, " ~ month_num")),
      data = d_fit, span = span
    ))
    p <- p %>% add_trace(
      data = d_fit, x = as.formula(paste0("~", x_col)), y = ~trend,
      name = name, type = "scatter", mode = "lines",
      legendgroup = legendgroup,
      line = list(color = color, width = 1.5, dash = "dash"),
      showlegend = FALSE, visible = visible, hoverinfo = "skip"
    )
  }
  p
}

plot_metric <- function(metric_col, y_label) {
  y_format <- switch(metric_col,
    rmse = ",.0f",
    mape = ".1f",
    ".3f"
  )
  shapes <- if (metric_col == "median_ratio") list(ratio_line) else list()

  sets <- list(
    list(data = total_test, label = "Test", color = "#2980b9"),
    list(data = total_train, label = "Train", color = "#e74c3c")
  )

  twps <- unique(metrics_test$township)
  twp_colors <- scales::hue_pal()(length(twps))

  p <- plot_ly()

  for (s in sets) {
    p <- add_metric_traces(
      p, s$data, metric_col, "month_year",
      name = paste0(s$label, " (total)"), color = s$color,
      legendgroup = s$label,
      line = list(color = s$color, width = 3, dash = "dot"),
      marker = list(
        color = s$color,
        size = 7, symbol = "diamond"
      )
    )
  }

  for (i in seq_along(twps)) {
    d <- filter(metrics_test, township == twps[i])
    p <- add_metric_traces(
      p, d, metric_col, "month_year",
      name = twps[i], color = twp_colors[i],
      legendgroup = twps[i], visible = "legendonly"
    )
  }

  p %>% plotly::layout(
    title = list(
      text = paste0(
        y_label, "<br><sup>Township traces show test-set data only</sup>"
      ),
      font = list(size = 14)
    ),
    xaxis = list(title = "Sale date", tickformat = "%b %Y"),
    yaxis = list(title = y_label, tickformat = y_format),
    hovermode = "x unified",
    legend = list(orientation = "v", x = 1.02, y = 1),
    shapes = shapes
  )
}

plot_vol_metric <- function(
    metric_col,
    y_label, loess_span = 0.75) {
  dollar <- metric_col == "rmse"
  y_format <- if (dollar) {
    "$,.0f"
  } else {
    switch(metric_col,
      mape = ".1f",
      ".3f"
    )
  }
  color <- "#2980b9"

  val_fmt <- if (dollar) {
    paste0("paste0('$',
    scales::comma(round(`", metric_col, "`, 0)))")
  } else {
    paste0("round(`", metric_col, "`, 3)")
  }

  df <- metrics_test %>%
    arrange(volume_pct) %>%
    mutate(trend = predict(loess(
      as.formula(paste0("`", metric_col, "` ~ volume_pct")),
      span = loess_span, na.action = na.exclude
    )))

  hover <- as.formula(paste0(
    "~paste0(township, '<br>',
    format(month_year, '%b %Y'),",
    "'<br>n = ', n,",
    "'<br>Volume: ', round(volume_pct, 2), '%',",
    "'<br>", y_label, ": ', ", val_fmt, ")"
  ))

  plot_ly() %>%
    add_trace(
      data = df, x = ~volume_pct,
      y = as.formula(paste0("~`", metric_col, "`")),
      name = "Test", size = ~n, sizes = c(5, 40),
      text = hover, hoverinfo = "text",
      type = "scatter", mode = "markers",
      marker = list(opacity = 0.6, color = color)
    ) %>%
    add_lines(
      data = df, x = ~volume_pct, y = ~trend,
      name = "Test (trend)",
      line = list(color = color, width = 2),
      showlegend = FALSE, hoverinfo = "skip"
    ) %>%
    plotly::layout(
      title = list(
        text = paste0(
          "Sales volume vs ", y_label, " by township-month (Test Set)"
        ),
        font = list(size = 14)
      ),
      xaxis = list(
        title = "Sales as % of township housing stock", tickformat = ".2f"
      ),
      yaxis = list(title = y_label, tickformat = y_format),
      legend = list(orientation = "h", x = 0, y = -0.2)
    )
}
```


## Metrics Over Time
::: {.panel-tabset}

### Median Ratio
```{r}
plot_metric("median_ratio", "Median Ratio")
```

### RMSE
```{r}
plot_metric("rmse", "RMSE ($)")
```

### MAPE
```{r}
plot_metric("mape", "MAPE (%)")
```

:::

## Volume vs. Metric
::: {.panel-tabset}

### Median Ratio
```{r}
plot_vol_metric("median_ratio", "Median Ratio")
```

### RMSE
```{r}
plot_vol_metric("rmse", "RMSE ($)")
```

### MAPE
```{r}
plot_vol_metric("mape", "MAPE (%)")
```

:::
