# Triad-Level Machine Learning Metrics

This section compares model performance metrics at the triad level between the
**training set** (the data the model learned from) and the **test set** (a
held-out sample the model never saw during training). Large gaps between train
and test metrics — particularly for RMSE, MDAPE, or MAE — may indicate
overfitting, where the model has memorized the training data rather than
learning generalizable patterns. The sale price and prediction distributions
are included to confirm that both splits are drawn from similar underlying
price ranges.

**Note**: Consistent with the rest of this report, while the model was trained on county-wide data, we report these splits at the triad level since that is where we are most concerned with overfitting.

```{r _train_test_setup_script}
source("../_setup.R")
```

```{r _train_test_pull_stats}
test_stats <- model_performance_test %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_stats <- model_performance_train %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_test_comparison_tbl <- dplyr::bind_rows(list(
  "train" = train_stats, "test" = test_stats
), .id = "split")
```


```{r _train_test_construct_table}
train_test_comparison_tbl %>%
  select(
    split, rmse, mdape, mae,
    sale_fmv_mean, sale_fmv_sd,
    estimate_fmv_mean, estimate_fmv_sd
  ) %>%
  mutate(across(c(
    rmse, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  ), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Train vs. Test Set Performance Metrics",
    col.names = c(
      "Split", "RMSE", "MdAPE", "MAE",
      "Mean Sale Price", "SD Sale Price",
      "Mean Estimate", "SD Estimate"
    )
  )
```

```{r _train_test_actual_price_distributions}
train_test_comparison_tbl %>%
  select(
    split,
    sale_fmv_min, sale_fmv_q25, sale_fmv_median, sale_fmv_q75, sale_fmv_max
  ) %>%
  mutate(across(starts_with("sale_fmv_"), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Actual Sale Price Distributions",
    col.names = c("Split", "Min", "Q1", "Median", "Q3", "Max")
  )
```

```{r _train_test_estimated_distributions}
train_test_comparison_tbl %>%
  select(
    split,
    estimate_fmv_min, estimate_fmv_q25, estimate_fmv_median,
    estimate_fmv_q75, estimate_fmv_max
  ) %>%
  mutate(across(starts_with("estimate_fmv_"), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Predicted Sale Price Distributions",
    col.names = c("Split", "Min", "Q1", "Median", "Q3", "Max")
  )
```
