# Triad-Level Machine Learning Metrics

This section compares model performance metrics at the triad level between the
**training set** (the data the model learned from) and the **test set** (a
held-out sample the model never saw during training). Large gaps between train
and test metrics — particularly for RMSE, MDAPE, or MAE — may indicate
overfitting, where the model has memorized the training data rather than
learning generalizable patterns. The sale price and prediction distributions
are included to confirm that both splits are drawn from similar underlying
price ranges.

**Note**: Consistent with the rest of this report, we report these splits at the triad level, even though the model was trained on county-wide data.  We do this, as we are most concerned with whether or not we have generalized appropriately (not-overfit) at the triad-level.

```{r _train_test_setup_script}
source("../_setup.R")
```

```{r}
test_stats <- model_performance_test %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_stats <- model_performance_train %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_test_comparison_tbl <- dplyr::bind_rows(list(
  "train" = train_stats, "test" = test_stats
), .id = "split")
```


```{r}
train_test_comparison_tbl %>%
  select(
    split, rmse, mdape, mae,
    sale_fmv_mean, sale_fmv_sd,
    estimate_fmv_mean, estimate_fmv_sd
  ) %>%
  mutate(across(c(
    rmse, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  ), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Train vs. Test Set Performance Metrics",
    col.names = c(
      "Split", "RMSE", "MdAPE", "MAE",
      "Mean Sale Price", "SD Sale Price",
      "Mean Estimate", "SD Estimate"
    )
  )
```

```{r}
train_test_comparison_tbl %>%
  select(
    split,
    sale_fmv_min, sale_fmv_q25, sale_fmv_median, sale_fmv_q75, sale_fmv_max
  ) %>%
  mutate(across(starts_with("sale_fmv_"), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Actual Sale Price Distributions",
    col.names = c("Split", "Min", "Q1", "Median", "Q3", "Max")
  )
```

```{r}
train_test_comparison_tbl %>%
  select(
    split,
    estimate_fmv_min, estimate_fmv_q25, estimate_fmv_median,
    estimate_fmv_q75, estimate_fmv_max
  ) %>%
  mutate(across(starts_with("estimate_fmv_"), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Predicted Sale Price Distributions",
    col.names = c("Split", "Min", "Q1", "Median", "Q3", "Max")
  )
```

```{r}
housing_stock <- assessment_pin %>%
  group_by(meta_township_code) %>%
  summarise(total_units = n_distinct(meta_pin), .groups = "drop") %>%
  mutate(meta_township_code = as.character(meta_township_code))

train_values <- training_data %>%
  select(
    meta_sale_price, meta_sale_date, meta_pin, sv_is_outlier,
    ind_pin_is_multicard
  ) %>%
  filter(sv_is_outlier == FALSE, ind_pin_is_multicard == FALSE) %>%
  left_join(
    assessment_card %>%
      group_by(meta_pin) %>%
      filter(n() == 1) %>%
      ungroup() %>%
      select(meta_pin, meta_township_code, pred_card_initial_fmv),
    by = "meta_pin"
  )

compute_metrics <- function(df, pred_col, units_col) {
  township_lookup <- df %>%
    distinct(meta_township_code) %>%
    mutate(
      meta_township_code = as.character(meta_township_code),
      township           = ccao::town_convert(meta_township_code)
    )

  df %>%
    mutate(meta_township_code = as.character(meta_township_code)) %>%
    left_join(housing_stock, by = "meta_township_code") %>%
    left_join(township_lookup, by = "meta_township_code") %>%
    mutate(
      .pred      = .data[[pred_col]],
      .units     = .data[[units_col]],
      ratio      = .pred / meta_sale_price,
      month_year = floor_date(as.Date(meta_sale_date), "month")
    ) %>%
    group_by(month_year, township, meta_township_code, .units) %>%
    summarise(
      median_ratio = median(ratio, na.rm = TRUE),
      rmse = sqrt(mean((.pred - meta_sale_price)^2, na.rm = TRUE)),
      mape = mean(abs((.pred - meta_sale_price) / meta_sale_price),
        na.rm = TRUE
      ) * 100,
      n = n(),
      .groups = "drop"
    ) %>%
    mutate(
      month_num  = as.numeric(month_year),
      volume_pct = (n / .units) * 100
    )
}

compute_total <- function(metrics) {
  metrics %>%
    group_by(month_year, month_num) %>%
    summarise(
      median_ratio = median(median_ratio, na.rm = TRUE),
      rmse         = sqrt(mean(rmse^2, na.rm = TRUE)),
      mape         = mean(mape, na.rm = TRUE),
      n            = sum(n),
      .groups      = "drop"
    )
}

compute_total_from_raw <- function(df, pred_col) {
  df %>%
    mutate(
      .pred      = .data[[pred_col]],
      ratio      = .pred / meta_sale_price,
      month_year = floor_date(as.Date(meta_sale_date), "month")
    ) %>%
    group_by(month_year) %>%
    summarise(
      median_ratio = median(ratio, na.rm = TRUE),
      rmse = sqrt(mean((.pred - meta_sale_price)^2, na.rm = TRUE)),
      mape = mean(abs((.pred - meta_sale_price) / meta_sale_price),
        na.rm = TRUE
      ) * 100,
      n = n(),
      .groups = "drop"
    ) %>%
    mutate(month_num = as.numeric(month_year))
}

metrics_test <- compute_metrics(test_card,
  pred_col = "pred_card_initial_fmv",
  units_col = "total_units"
)
total_train <- compute_total_from_raw(train_values,
  pred_col = "pred_card_initial_fmv"
)

sets <- list(
  list(
    metrics = metrics_test, total = compute_total(metrics_test),
    label = "Test", total_color = "#2980b9"
  ),
  list(
    metrics = NULL, total = total_train, label = "Train",
    total_color = "#e74c3c"
  )
)

# ── Helpers
add_trendline <- function(p, df, metric_col, color, name, visible = TRUE) {
  df_fit <- filter(df, !is.na(.data[[metric_col]]))
  if (nrow(df_fit) < 3) {
    return(p)
  }
  df_fit$trend <- predict(loess(
    as.formula(paste0(metric_col, " ~ month_num")),
    data = df_fit, span = 0.75
  ))
  add_trace(p,
    data = df_fit, x = ~month_year, y = ~trend,
    name = paste0(name, " (trend)"), type = "scatter", mode = "lines",
    line = list(color = color, width = 1.5, dash = "dash"),
    showlegend = FALSE, visible = visible, hoverinfo = "skip"
  )
}

toggle_trend_js <- "
  function(el, x) {
    el.on('plotly_legendclick', function(data) {
      var idx = data.curveNumber;
      if (idx % 2 !== 0) return false;
      var trendIdx = idx + 1;
      if (trendIdx >= el.data.length) return;
      var newVis = (el.data[idx].visible === true ||
      el.data[idx].visible === undefined) ? 'legendonly' : true;
      Plotly.restyle(el, { visible: newVis }, [trendIdx]);
    });
  }
"

ratio_line <- list(
  type = "line", xref = "paper", x0 = 0, x1 = 1,
  y0 = 1, y1 = 1, line = list(color = "red", width = 1.5, dash = "dash")
)

plot_metric <- function(metric_col, y_label, y_format = ".3f") {
  shapes <- if (metric_col == "median_ratio") list(ratio_line) else list()
  p <- plot_ly()

  for (s in sets) {
    p <- p %>%
      add_trace(
        data = s$total, x = ~month_year,
        y = as.formula(paste0("~", metric_col)),
        name = paste0(s$label, " (total)"), type = "scatter",
        mode = "lines+markers",
        line = list(color = s$total_color, width = 3, dash = "dot"),
        marker = list(color = s$total_color, size = 7, symbol = "diamond")
      ) %>%
      add_trendline(
        s$total, metric_col, s$total_color,
        paste0(s$label, " (total)")
      )

    # Township breakdown only for test set
    if (!is.null(s$metrics)) {
      twps <- unique(s$metrics$township)
      twp_colors <- scales::hue_pal()(length(twps))

      for (i in seq_along(twps)) {
        d <- filter(s$metrics, township == twps[i])
        p <- p %>%
          add_trace(
            data = d, x = ~month_year, y = as.formula(paste0("~", metric_col)),
            name = paste0(s$label, " - ", twps[i]), type = "scatter",
            mode = "lines+markers",
            line = list(color = twp_colors[i], width = 2),
            marker = list(color = twp_colors[i], size = 5),
            visible = "legendonly"
          ) %>%
          add_trendline(d, metric_col, twp_colors[i],
            paste0(
              s$label, " - ",
              twps[i]
            ),
            visible = "legendonly"
          )
      }
    }
  }

  p %>%
    plotly::layout(
      title     = list(text = y_label, font = list(size = 14)),
      xaxis     = list(title = "Sale date", tickformat = "%b %Y"),
      yaxis     = list(title = y_label, tickformat = y_format),
      hovermode = "x unified",
      legend    = list(orientation = "v", x = 1.02, y = 1),
      shapes    = shapes
    ) %>%
    htmlwidgets::onRender(toggle_trend_js)
}

plot_vol_metric <- function(metric_col,
                            y_label,
                            y_format = ".3f", dollar = FALSE) {
  p <- plot_ly()

  for (s in sets) {
    # Train has no township metrics, so skip the scatter/trend for it
    if (is.null(s$metrics)) next

    df <- s$metrics %>%
      arrange(volume_pct) %>%
      mutate(trend = predict(loess(
        as.formula(paste0(metric_col, " ~ volume_pct")),
        span = 0.75, na.action = na.exclude
      )))

    hover_val <- if (dollar) {
      paste0("~paste0(township,
      '<br>', format(month_year, '%b %Y'), '<br>n = ', n,
        '<br>Volume: ', round(volume_pct, 2), '%',
        '<br>", y_label, ": $', scales::comma(round(", metric_col, ", 0)))")
    } else {
      paste0("~paste0(township, '<br>',
      format(month_year, '%b %Y'), '<br>n = ', n,
        '<br>Volume: ', round(volume_pct, 2), '%',
        '<br>", y_label, ": ', round(", metric_col, ", 3))")
    }

    p <- p %>%
      add_trace(
        data = df,
        x = ~volume_pct, y = as.formula(paste0("~", metric_col)),
        name = s$label, legendgroup = s$label,
        size = ~n, sizes = c(5, 40),
        text = as.formula(hover_val), hoverinfo = "text",
        type = "scatter", mode = "markers",
        marker = list(opacity = 0.6, color = s$total_color)
      ) %>%
      add_lines(
        data = arrange(df, volume_pct),
        x = ~volume_pct, y = ~trend,
        name = s$label, legendgroup = s$label,
        line = list(color = s$total_color, width = 2),
        showlegend = FALSE, hoverinfo = "skip"
      )
  }

  p %>%
    plotly::layout(
      title = list(
        text = paste0(
          "Sales volume vs ",
          y_label, " by township-month"
        ),
        font = list(size = 14)
      ),
      xaxis = list(
        title = "Sales as % of township housing stock",
        tickformat = ".2f"
      ),
      yaxis = list(title = y_label, tickformat = y_format),
      legend = list(orientation = "h", x = 0, y = -0.2)
    )
}
```


## Metrics Over Time
::: {.panel-tabset}

### Median Ratio
```{r}
plot_metric("median_ratio", "Median Ratio", ".3f")
```

### RMSE
```{r}
plot_metric("rmse", "RMSE ($)", ",.0f")
```

### MAPE
```{r}
plot_metric("mape", "MAPE (%)", ".1f")
```

:::

## Volume vs. Metric
::: {.panel-tabset}

### Median Ratio
```{r}
plot_vol_metric("median_ratio", "Median Ratio", ".3f")
```

### RMSE
```{r}
plot_vol_metric("rmse", "RMSE ($)", ",.0f", dollar = TRUE)
```

### MAPE
```{r}
plot_vol_metric("mape", "MAPE (%)", ".1f")
```

:::
