# Triad-Level Machine Learning Metrics

This section compares model performance metrics at the triad level between the
**training set** (the data the model learned from) and the **test set** (a
held-out sample the model never saw during training). Large gaps between train
and test metrics — particularly for RMSE, MDAPE, or MAE — may indicate
overfitting, where the model has memorized the training data rather than
learning generalizable patterns. The sale price and prediction distributions
are included to confirm that both splits are drawn from similar underlying
price ranges.

**Note**: Consistent with the rest of this report, while the model was trained on county-wide data, we report these splits at the triad level since that is where we are most concerned with overfitting.

```{r _train_test_setup_script}
source("../_setup.R")
```

```{r _train_test_pull_stats}
test_stats <- model_performance_test %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_stats <- model_performance_train %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_test_comparison_tbl <- dplyr::bind_rows(list(
  "train" = train_stats, "test" = test_stats
), .id = "split")
```


```{r _train_test_construct_table}
train_test_comparison_tbl %>%
  select(
    split, rmse, mdape, mae,
    sale_fmv_mean, sale_fmv_sd,
    estimate_fmv_mean, estimate_fmv_sd
  ) %>%
  mutate(across(c(
    rmse, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  ), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Train vs. Test Set Performance Metrics",
    col.names = c(
      "Split", "RMSE", "MdAPE", "MAE",
      "Mean Sale Price", "SD Sale Price",
      "Mean Estimate", "SD Estimate"
    )
  )
```

```{r _train_test_actual_price_distributions}
train_test_comparison_tbl %>%
  select(
    split,
    sale_fmv_min, sale_fmv_q25, sale_fmv_median, sale_fmv_q75, sale_fmv_max
  ) %>%
  mutate(across(starts_with("sale_fmv_"), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Actual Sale Price Distributions",
    col.names = c("Split", "Min", "Q1", "Median", "Q3", "Max")
  )
```

```{r _train_test_estimated_distributions}
train_test_comparison_tbl %>%
  select(
    split,
    estimate_fmv_min, estimate_fmv_q25, estimate_fmv_median,
    estimate_fmv_q75, estimate_fmv_max
  ) %>%
  mutate(across(starts_with("estimate_fmv_"), scales::dollar)) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Predicted Sale Price Distributions",
    col.names = c("Split", "Min", "Q1", "Median", "Q3", "Max")
  )
```
housing_stock <- assessment_pin %>%
  group_by(meta_township_code) %>%
  summarise(total_units = n_distinct(meta_pin), .groups = "drop") %>%
  mutate(meta_township_code = as.character(meta_township_code))

compute_metrics <- function(df, pred_col) {
  df %>%
    mutate(
      meta_township_code = as.character(meta_township_code),
      township           = ccao::town_convert(meta_township_code),
      .pred              = .data[[pred_col]],
      ratio              = .pred / meta_sale_price,
      month_year         = floor_date(as.Date(meta_sale_date), "month")
    ) %>%
    left_join(housing_stock, by = "meta_township_code") %>%
    group_by(month_year, township, meta_township_code, total_units) %>%
    summarise(
      median_ratio = median(ratio, na.rm = TRUE),
      rmse = sqrt(mean((.pred - meta_sale_price)^2, na.rm = TRUE)),
      mape = mean(abs((.pred - meta_sale_price) / meta_sale_price),
        na.rm = TRUE
      ) * 100,
      n = n(),
      .groups = "drop"
    ) %>%
    mutate(
      month_num  = as.numeric(month_year),
      volume_pct = (n / total_units) * 100
    )
}

compute_total_from_raw <- function(df, pred_col) {
  df %>%
    mutate(
      .pred      = .data[[pred_col]],
      ratio      = .pred / meta_sale_price,
      month_year = floor_date(as.Date(meta_sale_date), "month")
    ) %>%
    group_by(month_year) %>%
    summarise(
      median_ratio = median(ratio, na.rm = TRUE),
      rmse = sqrt(mean((.pred - meta_sale_price)^2, na.rm = TRUE)),
      mape = mean(abs((.pred - meta_sale_price) / meta_sale_price),
        na.rm = TRUE
      ) * 100,
      n = n(),
      .groups = "drop"
    ) %>%
    mutate(month_num = as.numeric(month_year))
}

metrics_test <- compute_metrics(test_card,
  pred_col = "pred_card_initial_fmv"
)
total_test <- compute_total_from_raw(test_card,
  pred_col = "pred_card_initial_fmv"
)
total_train <- compute_total_from_raw(train_card,
  pred_col = "pred_card_initial_fmv"
)

ratio_line <- list(
  type = "line", xref = "paper", x0 = 0, x1 = 1,
  y0 = 1, y1 = 1, line = list(color = "red", width = 1.5, dash = "dash")
)

plot_metric <- function(metric_col, y_label) {
  y_format <- switch(metric_col,
    rmse = ",.0f",
    mape = ".1f",
    ".3f"
  )
  shapes <- if (metric_col == "median_ratio") list(ratio_line) else list()

  sets <- list(
    list(data = total_test, label = "Test", color = "#2980b9"),
    list(data = total_train, label = "Train", color = "#e74c3c")
  )

  twps <- unique(metrics_test$township)
  twp_colors <- scales::hue_pal()(length(twps))

  p <- plot_ly()
  for (s in sets) {
    df_fit <- filter(s$data, !is.na(.data[[metric_col]]))
    if (nrow(df_fit) >= 3) {
      df_fit$trend <- predict(loess(
        as.formula(paste0(metric_col, " ~ month_num")),
        data = df_fit, span = 0.75
      ))
    }
    p <- p %>%
      add_trace(
        data = s$data, x = ~month_year,
        y = as.formula(paste0("~", metric_col)),
        name = paste0(s$label, " (total)"), type = "scatter",
        mode = "lines+markers",
        legendgroup = s$label,
        line = list(color = s$color, width = 3, dash = "dot"),
        marker = list(color = s$color, size = 7, symbol = "diamond")
      )
    if (nrow(df_fit) >= 3) {
      p <- p %>% add_trace(
        data = df_fit, x = ~month_year, y = ~trend,
        name = paste0(s$label, " (trend)"), type = "scatter", mode = "lines",
        legendgroup = s$label,
        line = list(color = s$color, width = 1.5, dash = "dash"),
        showlegend = FALSE, hoverinfo = "skip"
      )
    }
  }

  for (i in seq_along(twps)) {
    d <- filter(metrics_test, township == twps[i])
    d_fit <- filter(d, !is.na(.data[[metric_col]]))
    if (nrow(d_fit) >= 3) {
      d_fit$trend <- predict(loess(
        as.formula(paste0(metric_col, " ~ month_num")),
        data = d_fit, span = 0.75
      ))
    }
    p <- p %>%
      add_trace(
        data = d, x = ~month_year, y = as.formula(paste0("~", metric_col)),
        name = twps[i], type = "scatter", mode = "lines+markers",
        legendgroup = twps[i],
        line = list(color = twp_colors[i], width = 2),
        marker = list(color = twp_colors[i], size = 5),
        visible = "legendonly"
      )
    if (nrow(d_fit) >= 3) {
      p <- p %>% add_trace(
        data = d_fit, x = ~month_year, y = ~trend,
        name = twps[i], type = "scatter", mode = "lines",
        legendgroup = twps[i],
        line = list(color = twp_colors[i], width = 1, dash = "dash"),
        showlegend = FALSE, visible = "legendonly", hoverinfo = "skip"
      )
    }
  }

  p %>% plotly::layout(
    title     = list(text = y_label, font = list(size = 14)),
    xaxis     = list(title = "Sale date", tickformat = "%b %Y"),
    yaxis     = list(title = y_label, tickformat = y_format),
    hovermode = "x unified",
    legend    = list(orientation = "v", x = 1.02, y = 1),
    shapes    = shapes
  )
}

plot_vol_metric <- function(
    metric_col, y_label,
    dollar = FALSE, loess_span = 0.75) {
  y_format <- if (dollar) {
    "$,.0f"
  } else {
    switch(metric_col,
      mape = ".1f",
      ".3f"
    )
  }
  color <- "#2980b9"

  df <- metrics_test %>%
    arrange(volume_pct) %>%
    mutate(trend = predict(loess(
      as.formula(paste0("`", metric_col, "` ~ volume_pct")),
      span = loess_span, na.action = na.exclude
    )))

  hover <- if (dollar) {
    as.formula(paste0(
      "~paste0(township, '<br>', format(month_year, '%b %Y'),
      '<br>n = ', n,
      '<br>Volume: ', round(volume_pct, 2), '%',
      '<br>", y_label, ": $', scales::comma(round(`", metric_col, "`, 0)))"
    ))
  } else {
    as.formula(paste0(
      "~paste0(township, '<br>', format(month_year, '%b %Y'),
      '<br>n = ', n,
      '<br>Volume: ', round(volume_pct, 2), '%',
      '<br>", y_label, ": ', round(`", metric_col, "`, 3))"
    ))
  }

  plot_ly() %>%
    add_trace(
      data = df, x = ~volume_pct,
      y = as.formula(paste0("~`", metric_col, "`")),
      name = "Test", size = ~n, sizes = c(5, 40),
      text = hover, hoverinfo = "text",
      type = "scatter", mode = "markers",
      marker = list(opacity = 0.6, color = color)
    ) %>%
    add_lines(
      data = df, x = ~volume_pct, y = ~trend,
      name = "Test (trend)",
      line = list(color = color, width = 2),
      showlegend = FALSE, hoverinfo = "skip"
    ) %>%
    plotly::layout(
      title = list(
        text = paste0("Sales volume vs ", y_label, " by township-month"),
        font = list(size = 14)
      ),
      xaxis = list(
        title = "Sales as % of township housing stock",
        tickformat = ".2f"
      ),
      yaxis = list(title = y_label, tickformat = y_format),
      legend = list(orientation = "h", x = 0, y = -0.2)
    )
}
```


## Metrics Over Time
::: {.panel-tabset}

### Median Ratio
```{r}
plot_metric("median_ratio", "Median Ratio")
```

### RMSE
```{r}
plot_metric("rmse", "RMSE ($)")
```

### MAPE
```{r}
plot_metric("mape", "MAPE (%)")
```

:::

## Volume vs. Metric
::: {.panel-tabset}

### Median Ratio
```{r}
plot_vol_metric("median_ratio", "Median Ratio")
```

### RMSE
```{r}
plot_vol_metric("rmse", "RMSE ($)", dollar = TRUE)
```

### MAPE
```{r}
plot_vol_metric("mape", "MAPE (%)")
```

:::
