# Triad-Level Machine Learning Metrics

This section compares model performance metrics at the triad level between the
**training set** (the data the model learned from) and the **test set** (a
held-out sample the model never saw during training). Large gaps between train
and test metrics — particularly for RMSE, MDAPE, or MAE — may indicate
overfitting, where the model has memorized the training data rather than
learning generalizable patterns. The sale price and prediction distributions
are included to confirm that both splits are drawn from similar underlying
price ranges.

```{r _train_test_setup_script}
source("../_setup.R")

if (!exists("model_performance_train")) {
  model_performance_train <-
    arrow::read_parquet(paths$output$performance_train$local)
}
```

```{r}
test_stats <- model_performance_test %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_stats <- model_performance_train %>%
  filter(
    geography_type == "triad_code",
    triad_code == run_triad_code,
    by_class == FALSE
  ) %>%
  select(
    rmse, mdape, mae, starts_with("sale_fmv_"),
    starts_with("estimate_fmv_")
  )

train_test_comparison_tbl <- dplyr::bind_rows(list(
  "train" = train_stats, "test" = test_stats
), .id = "split")
```


```{r}
train_test_comparison_tbl %>%
  select(
    split, rmse, mdape, mae,
    sale_fmv_mean, sale_fmv_sd,
    estimate_fmv_mean, estimate_fmv_median
  ) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Train vs. Test Set Performance Metrics"
  )
```

```{r}
train_test_comparison_tbl %>%
  select(
    split,
    sale_fmv_min, sale_fmv_q25, sale_fmv_median, sale_fmv_q75, sale_fmv_max
  ) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Actual Sale Price Distributions, for Reference"
  )
```

```{r}
train_test_comparison_tbl %>%
  select(
    split,
    estimate_fmv_min, estimate_fmv_q25, estimate_fmv_median,
    estimate_fmv_q75, estimate_fmv_max
  ) %>%
  kable(
    digits = 3,
    format.args = list(big.mark = ","),
    caption = "Predicted Value Distributions, for Reference"
  )
```
